### 2016-04-18 sig-alternate-sample.tex
initial empty template

% This is "sig-alternate.tex" V2.1 April 2013
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate-05-2015}
\usepackage{xcolor}
\newcommand\todo[1]{\textcolor{red}{#1}}

\begin{document}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{10.475/123_4}

% ISBN
\isbn{123-4567-24-567/08/06}

%Conference
\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}

\acmPrice{\$15.00}

%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Toward Automatic Translation from OpenACC to OpenMP 4}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
\alignauthor
G.K.M. Tobin\titlenote{The secretary disavows
any knowledge of this author's actions.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{P.O. Box 1212}\\
       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{webmaster@marysville-ohio.com}
% 3rd. author
\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
one who did all the really hard work.}\\
       \affaddr{The Th{\o}rv{\"a}ld Group}\\
       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       \affaddr{Hekla, Iceland}\\
       \email{larst@affiliation.org}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Lawrence P. Leipuner\\
       \affaddr{Brookhaven Laboratories}\\
       \affaddr{Brookhaven National Lab}\\
       \affaddr{P.O. Box 5000}\\
       \email{lleipuner@researchlabs.org}
% 5th. author
\alignauthor Sean Fogarty\\
       \affaddr{NASA Ames Research Center}\\
       \affaddr{Moffett Field}\\
       \affaddr{California 94035}\\
       \email{fogartys@amesres.org}
% 6th. author
\alignauthor Charles Palmer\\
       \affaddr{Palmer Research Laboratories}\\
       \affaddr{8600 Datapoint Drive}\\
       \affaddr{San Antonio, Texas 78229}\\
       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}

\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%
% End generated code
%

%
%  Use this command to print the description
%
\printccsdesc

% We no longer use \terms command
%\terms{Theory}

\keywords{ACM proceedings; \LaTeX; text tagging}

\section{Introduction}
\todo{1 page - introduction/summary of the work - Dr. Overbey}

\section{Motivation}
\subsection{History of OpenACC and OpenMP}
\todo{Alexander}
\subsection{Future of OpenACC and OpenMP}
\todo{Alexander}
\subsection{Relevance to XSEDE}
\todo{Dr. Overbey}

\section{Overview}
\subsection{OpenACC Directives}
\todo{Alexander}
\subsection{OpenMP 4 Directives}
\todo{Alexander}

\section{Translation}
\subsection{Tool for Translation}
\todo{1 page including screenshot(s) - description of our tool: what directives it handles, how it works, where to get it - Nawrin}
\subsection{Algorithm}
\todo{0.5 page - description of the conversion algorithm (what clauses translate to what?) - Nawrin/Dr.Overbey} 
\section{Evaluation}
\subsection{Test Corpus}
\subsection{Performance}
\todo{1 page including tables - evaluation: what code did we test it on; how many lines changed; what performance numbers did we get - Nawrin}

\section{Limitations and Future Work}
\todo{1 page - limitations, including directives not handled, and future work - Dr. Overbey}

\section{Related Work}
\section{Conclusion}
\todo{Dr. Overbey}

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
%\appendix
%Appendix A
%\section{Headings in Appendices}
%The rules about hierarchical headings discussed above for
%the body of the article are different in the appendices.
%In the \textbf{appendix} environment, the command
%\textbf{section} is used to
%indicate the start of each Appendix, with alphabetic order
%designation (i.e. the first is A, the second B, etc.) and
%a title (if you include one).  So, if you need
%hierarchical structure
%\textit{within} an Appendix, start with \textbf{subsection} as the
%highest level. Here is an outline of the body of this
%document in Appendix-appropriate form:
%\subsection{Introduction}
%\subsection{The Body of the Paper}
%\subsubsection{Type Changes and  Special Characters}
%\subsubsection{Math Equations}
%\paragraph{Inline (In-text) Equations}
%\paragraph{Display Equations}
%\subsubsection{Citations}
%\subsubsection{Tables}
%\subsubsection{Figures}
%\subsubsection{Theorem-like Constructs}
%\subsubsection*{A Caveat for the \TeX\ Expert}
%\subsection{Conclusions}
%\subsection{Acknowledgments}
%\subsection{Additional Authors}
%This section is inserted by \LaTeX; you do not insert it.
%You just add the names and information in the
%\texttt{{\char'134}additionalauthors} command at the start
%of the document.
%\subsection{References}
%Generated by bibtex from your ~.bib file.  Run latex,
%then bibtex, then latex twice (to resolve references)
%to create the ~.bbl file.  Insert that ~.bbl file into
%the .tex source file and comment out
%the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
%\section{More Help for the Hardy}
%The sig-alternate.cls file itself is chock-full of succinct
%and helpful comments.  If you consider yourself a moderately
%experienced to expert user of \LaTeX, you may find reading
%it useful but please remember not to change it.
%\balancecolumns % GM June 2007
% That's all folks!
\end{document}

### 2016-04-20 sig-alternate-sample.tex
added draft

@@
 \usepackage{xcolor}
-\newcommand\todo[1]{\textcolor{red}{#1}}
+\usepackage{xspace}
+\newcommand\todo[1]{\textcolor{red}{** #1}}
+\newcommand{\pooda}{\textsc{POOD}\xspace}
+
+\newcommand{\pood}{\textsc{Porting OpenACC To OpenMP Directives}\xspace}
+
+\newcommand{\squishlist}{
+    \begin{list}{$\bullet$}
+     { \setlength{\itemsep}{0pt}      \setlength{\parsep}{3pt}
+       \setlength{\topsep}{3pt}       \setlength{\partopsep}{0pt}
+       \setlength{\leftmargin}{1.5em} \setlength{\labelwidth}{1em}
+       \setlength{\labelsep}{0.5em} } }
 
@@
 \section{Translation}
-\subsection{Tool for Translation}
 \todo{1 page including screenshot(s) - description of our tool: what directives it handles, how it works, where to get it - Nawrin}
+The major compilers supporting OpenACC are Portland Group (PGI), Cray, and CAPS.
+On the other hand, Intel has decided not to support OpenACC until accelerators are fully integrated into OpenMP.
+We describe an automated approach to assist programmers in converting OpenACC to OpenMP, so that same syntax could be used for both Cray and Intel compiler while maintaining the performance.
+We describe a refactoring (POOD) for this purpose.
+
+\subsection{Translation Tool}
+We prototyped the refactoring in a PLDT OpenACC Eclipse plug-in (C analyses and refactorings for OpenACC).
+Our tool support various data flow analyses, dependence analysis, refactorings, and parser for OpenACC directives.
+It provides an user interface (UI) for refactorings.
+Other than UI, it also supports command line interface (CLI), so one can run the refactorings outside Eclipse.
+
+Our tool does not handle all the OpenACC directives.
+We transformed the directives present in EPCC level 1 OpenACC benchmark except ``kernel" directive.
+When user invoke a refactoring on a C source file, refactoring engine performs the precondition checking.
+If preconditions are violated, it shows appropriate error message to user.
+Then analyses and source translations are executed.
+
+\subsection{Porting OpenACC To OpenMP Directives (POOD)}
+The \pood refactoring converts OpenACC directives to OpenMP 4.0 directives.
+
+\textbf{Motivation:} While porting OpenACC to OpenMP it is necessary for directives to be compatible. Without automated tooling, a programmer has to manually transform each directives.
+
+\textbf{Precondition:} A programmer selects a C source file and invokes the \pooda transformation. The following precondition is checked:
+
+\begin{itemize}
+\item Source file contains OpenACC directives.
+\item OpenACC directives in source code supported by our transformation tool.
+\end{itemize}
+
+\textbf{Mechanism:} 
 \subsection{Algorithm}
@@
 \section{Evaluation}
+\todo{1 page including tables - evaluation: what code did we test it on; how many lines changed; what performance numbers did we get - Nawrin}
 \subsection{Test Corpus}
 \subsection{Performance}
-\todo{1 page including tables - evaluation: what code did we test it on; how many lines changed; what performance numbers did we get - Nawrin}
 
@@
 %\appendix
-%Appendix A
-%\section{Headings in Appendices}
-%The rules about hierarchical headings discussed above for
-%the body of the article are different in the appendices.
-%In the \textbf{appendix} environment, the command
-%\textbf{section} is used to
-%indicate the start of each Appendix, with alphabetic order
-%designation (i.e. the first is A, the second B, etc.) and
-%a title (if you include one).  So, if you need
-%hierarchical structure
-%\textit{within} an Appendix, start with \textbf{subsection} as the
-%highest level. Here is an outline of the body of this
-%document in Appendix-appropriate form:
-%\subsection{Introduction}
-%\subsection{The Body of the Paper}
-%\subsubsection{Type Changes and  Special Characters}
-%\subsubsection{Math Equations}
-%\paragraph{Inline (In-text) Equations}
-%\paragraph{Display Equations}
-%\subsubsection{Citations}
-%\subsubsection{Tables}
-%\subsubsection{Figures}
-%\subsubsection{Theorem-like Constructs}
-%\subsubsection*{A Caveat for the \TeX\ Expert}
-%\subsection{Conclusions}
-%\subsection{Acknowledgments}
-%\subsection{Additional Authors}
-%This section is inserted by \LaTeX; you do not insert it.
-%You just add the names and information in the
-%\texttt{{\char'134}additionalauthors} command at the start
-%of the document.
-%\subsection{References}
-%Generated by bibtex from your ~.bib file.  Run latex,
-%then bibtex, then latex twice (to resolve references)
-%to create the ~.bbl file.  Insert that ~.bbl file into
-%the .tex source file and comment out
-%the command \texttt{{\char'134}thebibliography}.
-% This next section command marks the start of
-% Appendix B, and does not continue the present hierarchy
-%\section{More Help for the Hardy}
-%The sig-alternate.cls file itself is chock-full of succinct
-%and helpful comments.  If you consider yourself a moderately
-%experienced to expert user of \LaTeX, you may find reading
-%it useful but please remember not to change it.
-%\balancecolumns % GM June 2007
-% That's all folks!
 \end{document}

### 2016-04-21 sig-alternate-sample.tex
added draft

@@
 
-\textbf{Mechanism:} 
-\subsection{Algorithm}
+\textbf{Mechanism:} The refactoring introduces OpenMP directives that correspond to OpenACC directives.
+
+POOD first determines all the preprocessor include statements and check whether there is any ``openacc.h".
+If openacc header presents POOD removes it.\todo{need to remove openacc functions}.
+From the precondition checking we have the OpenACC preprocessor pragma statements that we want to transform.
+For each directives Abstract Syntax Tree gives the exact position.
+POOD removes the existing pragma statment and inserts OpenMP directives on that position based on the directive clause.
+Table \ref{table:1} lists the OpenACC and corresponding OpenMP clauses that POOD transforms.
+For any OpenACC data directives with copy, copyin, or copyout clause POOD performs NameBinding analyses\todo{fix} for the variables to find out their declaration.
+If these variables are declared globally then POOD changes them to local.
+When all the transformations have completed, POOD writes the changes to file.
+
+\begin{table}[t]
+	\begin{minipage}{0.47\textwidth}
+%\small
+\caption{Equivqlent OpenACC and OpenMP clause}
+\centering
+\begin{tabular}{ | l@{\hspace{0.1cm}}|  @{\hspace{0.4cm}}r@{\hspace{0.1cm}}|} 
+ \hline
+ \hspace*{0.4cm} OpenACC Clause & OpenMP Clause\hspace*{0.6cm}\\
+ \hspace*{0.5cm} (\#pragma acc)%\footnote{The ID is used to refer to each application in the rest of the paper.} 
+  & (\#pragma omp)\hspace*{0.6cm} \\
+ \hline
+ data & target data \\ 
+ copy(A[0:n*n]) & map(tofrom:A[0:n*n]) \\
+ copyin(A[0:n*n]) & map(to:A[0:n*n]) \\
+ copyout(A[0:n*n]) & map(from:A[0:n*n]) \\
+ create(A[0:n*n]) & map(to:A[0:n*n]) \\ 
+ parallel & target/teams \\ 
+ parallel loop private(t1,t2) & parallel for private(t1,t2) \\
+ loop & for \\
+ loop private(tmp) & for private(tmp) \\
+ loop reduction(+:t1,t2) & for reduction(+:t1,t2) \\
+ loop independent & for independent \\
+ parallel loop & target/teams/distribute \\
+ \hline
+\end{tabular}
+
+\label{table:1}
+\end{minipage}
+\end{table}
+%\normalsize
+
+\textbf{Example:}
+
 \todo{0.5 page - description of the conversion algorithm (what clauses translate to what?) - Nawrin/Dr.Overbey} 

### 2016-04-21 sig-alternate-sample.tex
Added grammar and translation rules

@@
 
+\begin{figure}
+\begin{tabular}{rcl}
+\textit{acc-directive} &$\rightarrow$& \texttt{\#pragma acc data\ }\textit{data-clauses} \\
+                       &     $|$     & \texttt{\#pragma acc parallel loop} \\
+                       &     $|$     & \texttt{\#pragma acc parallel} \\
+                       &     $|$     & \texttt{\#pragma acc loop\ }\textit{loop-clauses} \\
+\textit{data-clause}   &$\rightarrow$& \texttt{copyin(}\textit{vars}\texttt{)} \\
+                       &     $|$     & \texttt{copyout(}\textit{vars}\texttt{)} \\
+                       &     $|$     & \texttt{create(}\textit{vars}\texttt{)} \\
+\textit{loop-clause}   &$\rightarrow$& \texttt{independent} \\
+                       &     $|$     & \texttt{private(}\textit{vars}\texttt{)} \\
+                       &     $|$     & \texttt{reduction(}\textit{reductions}\texttt{)} \\
+\end{tabular}
+\caption{OpenACC subset grammar accepted by the translation tool.  The syntax
+of \textit{vars} and \textit{reductions} is omitted.  The nonterminals
+\textit{data-clauses} and \textit{loop-clauses} correspond to sequences of zero
+or more phrases derived from \textit{data-clause} and \textit{loop-clause},
+respectively.}
+\label{fig:grammar}
+\end{figure}
+
+\newcommand{\translate}[1]{\ensuremath{\mathcal{T}[\![ #1 ]\!]}}
+\begin{figure*}
+\begin{align*}
+%
+\translate{\texttt{\#pragma acc data\ }\textit{data-clauses}} &=
+   \texttt{\#pragma omp target data } \translate{\textit{data-clauses}} \\
+%
+\translate{\texttt{copyin(}\textit{vars}\texttt{)}} &=
+   \texttt{map(to:}\textit{vars}\texttt{)} \\
+%
+\translate{\texttt{copyout(}\textit{vars}\texttt{)}} &=
+   \texttt{map(from:}\textit{vars}\texttt{)} \\
+%
+\translate{\texttt{create(}\textit{vars}\texttt{)}} &=
+   \texttt{map(alloc:}\textit{vars}\texttt{)} \\
+%
+\translate{\texttt{\#pragma acc parallel loop}} &=
+   \left\{\parbox{2in}{%
+   \texttt{\#pragma omp target} \\
+   \texttt{\#pragma omp teams} \\
+   \texttt{\#pragma omp distribute}}\right. \\
+%
+\translate{\texttt{\#pragma acc parallel}} &= %\textit{parallel-clauses}} &=
+   \left\{\parbox{2in}{%
+   \texttt{\#pragma omp target} \\
+   \texttt{\#pragma omp teams}}\right. \\
+%
+\translate{\texttt{\#pragma acc loop\ }\textit{loop-clauses}} &=
+   \texttt{\#pragma omp parallel for\ }\textit{loop-clauses} \\
+%
+\end{align*}
+\caption{Translation rules from OpenACC to OpenMP~4.  The function
+$\mathcal{T}$ maps an OpenACC directive to a sequence of one or more OpenMP
+directives.}
+\label{fig:rules}
+\end{figure*}
+
 \textbf{Example:}

### 2016-04-24 sig-alternate-sample.tex
Added grammar and translation rules

--- 2016-04-21 sig-alternate-sample.tex
@@
 
+\subsection{A Subset of OpenACC}
+For the purposes of prototyping, we designed our tool to accept the subset of
+OpenACC described by the grammar in Figure~\ref{fig:grammar}.  This includes
+all of the directives and clauses used by the EPCC Benchmarks, as well as the
+other example programs in our test suite.  While it omits some important
+features---asynchronous execution, unstructured data lifetimes, etc.---it does
+correspond to a useful, commonly-used subset of OpenACC.
+
+\begin{figure}
+\begin{tabular}{rcl}
+\textit{acc-directive} &$\rightarrow$& \texttt{\#pragma acc data\ }\textit{data-clauses} \\
+                       &     $|$     & \texttt{\#pragma acc parallel}\texttt{\ }\textit{parallel-clauses} \\
+                       &     $|$     & \texttt{\#pragma acc loop\ }\textit{type}\texttt{\ }\textit{loop-clauses} \\
+\textit{type}          &$\rightarrow$& \texttt{gang} \\
+                       &     $|$     & \texttt{vector} \\
+                       &     $|$     & \texttt{seq} \\
+                       &     $|$     & $\epsilon$ \\
+\textit{data-clause}   &$\rightarrow$& \texttt{copyin(}\textit{vars}\texttt{)} \\
+                       &     $|$     & \texttt{copyout(}\textit{vars}\texttt{)} \\
+                       &     $|$     & \texttt{copy(}\textit{vars}\texttt{)} \\
+                       &     $|$     & \texttt{create(}\textit{vars}\texttt{)} \\
+\textit{parallel-clause}&$\rightarrow$&\texttt{num\_gangs(}\textit{n}\texttt{)} \\
+                       &     $|$     & \texttt{vector\_length(}\textit{n}\texttt{)} \\
+                       &     $|$     & \texttt{private(}\textit{vars}\texttt{)} \\
+                       &     $|$     & \texttt{reduction(}\textit{reductions}\texttt{)} \\
+\textit{loop-clause}   &$\rightarrow$& \texttt{independent} \\
+                       &     $|$     & \texttt{private(}\textit{vars}\texttt{)} \\
+\end{tabular}
+\caption{OpenACC subset grammar accepted by the translation tool.  Syntax is
+omitted for \textit{vars}, \textit{reductions}, and integers \textit{n}.  The
+nonterminals \textit{data-clauses} and \textit{loop-clauses} correspond to
+sequences of zero or more phrases derived from \textit{data-clause} and
+\textit{loop-clause}, respectively.  In addition to the directives shown,
+\texttt{\#pragma acc parallel loop} is supported, and treated as syntactic
+sugar for a \texttt{loop} directive nested directly under a \texttt{parallel}
+directive.}
+\label{fig:grammar}
+\end{figure}
+
+\subsection{Considerations in Translation}
+In theory, the goals of the present paper are modest: We are interested in
+translating a subset of OpenACC (Figure~\ref{fig:grammar}) to OpenMP~4,
+targeting the same accelerator devices (NVIDIA GPUs).  In theory, the two APIs
+are very closely aligned, and the translation should be straightforward.
+
+In practice, the situation is not so simple.  The OpenACC and OpenMP
+specifications define an \emph{abstract} execution model.  They do not dictate
+how that model is mapped to specific hardware devices.  This is left to the
+compiler.  Our OpenACC programs were compiled using PGCC~16.1, the standard
+commercial compiler for OpenACC\@.  We compiled OpenMP~4 using a custom build
+of Clang~3.8\footnote{See \url{https://parallel-computing.pro/index.php/9-cuda/43-openmp-4-0-on-nvidia-cuda-gpus}} (PGCC does not yet support
+OpenMP~4\footnote{Cray's compilers support both, but there are currently no
+Cray resources in the XSEDE ecosystem.}).  As we will show, directives that
+appear similar may be translated quite differently by the two compilers.
+
+\subsubsection{Work Distribution Differs}
+
+\begin{figure*}
+\begin{center}
+\begin{minipage}{4in}
+{\footnotesize
+\begin{verbatim}
+#pragma acc data copyin(A[0:N], B[0:N]), copyout(C[0:N])
+#pragma omp target data map(to:A[0:N], B[0:N]), map(from:C[0:N])
+{
+    #pragma acc parallel loop
+    #pragma omp target teams distribute parallel for
+    for (int i = 0; i < N; ++i)
+    {
+        C[i] = A[i] + B[i];
+    }
+}
+\end{verbatim}
+}
+\end{minipage}
+\end{center}
+\caption{Vector addition}
+\label{fig:vecadd}
+\end{figure*}
+
+Perhaps the simplest example of directive-based accelerator code is the vector
+addition kernel shown in Figure~\ref{fig:vecadd}.  It adds two vectors $A$ and
+$B$ on an accelerator device, storing the sum in $C$.  While the directives
+direct the compiler to parallelize the loop on the accelerator, the compiler
+can decide (in this case) how to divide the loop iterations among threads
+and thread blocks on the CUDA device.  When $N = 10000$:
+\begin{itemize}
+\item PGCC generates a grid of 79 thread blocks with 128 threads per block;
+kernel execution time is about $3~\mu s$ on a Tesla~K40.
+\item Clang generates a grid of 15 thread blocks with 1024 threads per block;
+kernel execution time is about $30~\mu s$ on a Tesla~K40.\footnote{Adding
+\texttt{num\_teams(79) thread\_limit(128)} to the OpenMP code (to match the
+grid and block dimensions of the OpenACC code) actually increases the kernel
+execution time to about $80~\mu s$.
+%Adding \texttt{num\_gangs(15) vector\_length(1024)} to the OpenACC code does
+%not change the performance significantly.
+}
+\end{itemize}
+
+It is also worth noting that PGCC generates code where all threads are active
+in all thread blocks except the last (presumably to minimize control
+divergence).  Clang generates code that assigns an approximately equal number
+of threads to each thread block, so inactive threads may be present in
+several blocks.
+
+\begin{figure*}
+\begin{minipage}{.5\textwidth}
+{\footnotesize
+\begin{verbatim}
+// OpenACC
+#pragma acc parallel loop
+for (int i = 0; i < 10; i++) {      // Gang
+    for (int j = 0; j < 10; j++) {  // Vector
+        ...
+    }
+}
+
+\end{verbatim}
+}
+\end{minipage}
+\begin{minipage}{.5\textwidth}
+{\footnotesize
+\begin{verbatim}
+// OpenMP 4 Device Directives
+#pragma omp target teams distribute
+for (int i = 0; i < 10; i++) {
+    #pragma omp parallel for
+    for (int j = 0; j < 10; j++) {
+        ...
+    }
+}
+\end{verbatim}
+}
+\end{minipage}
+\caption{Doubly nested loops.}
+\label{fig:nest}
+\end{figure*}
+
+More important is the handling of loop nests, such as the simple
+doubly nested loops in Figure~\ref{fig:nest}.
+
+In OpenACC, the compiler has some freedom to decide how to parallelize the loop
+nest.  PGCC will typically run the outermost loop as a gang loop and the
+innermost loop as a vector loop.  Given the OpenACC code in
+Figure~\ref{fig:nest}, the compiler does exactly this: the iterations of the
+$i$-loop are divided among the thread blocks, and the iterations of the
+$j$-loop are divided among the threads within each block.
+
+In OpenMP, work distribution is not so automatic.  Consider these
+``obvious'' but incorrect translations of the OpenACC code into OpenMP: The
+\texttt{\#pragma acc parallel loop} directive could be replaced by
+\begin{itemize}
+\item \texttt{\#pragma omp target teams}.  In this case, the entire loop
+nest would be run redundantly by all thread blocks.
+\item \texttt{\#pragma omp target teams distribute}.  In this case, the
+iterations of the $i$-loop would be divided among thread blocks, but only
+one thread would be active within each block.
+\item \texttt{\#pragma omp target teams distribute parallel for}.  Here,
+the iterations of the $i$-loop are distributed among threads and thread blocks.
+\end{itemize}
+In all three cases, \emph{the inner $j$-loop runs sequentially}, because the
+parallelization directive is applied to only one loop.  OpenMP requires the
+programmer to explicitly specify how to run each loop in the nest.  The
+translation most similar to the OpenACC code is shown on the right in
+Figure~\ref{fig:nest}: The iterations of the $i$-loop are divided among thread
+blocks using the \texttt{distribute} directive, and the iterations of the
+$j$-loop are divided among threads in the block using the \texttt{parallel for}
+directive.
+
+%\subsubsection{Intrinsics Differ}
+
+%\subsubsection{Reductions Differ}
+
+\subsubsection{Data Transfer May Differ}
+
+In both OpenACC and OpenMP, if \texttt{data}/\texttt{target data} directives
+are omitted, the compiler will automatically determine what data needs to be
+transferred between the host and the accelerator.  This is another case where
+compilers may differ.  In the vector addition example
+(Figure~\ref{fig:vecadd}), consider what happens when $A$, $B$, and $C$ are
+\emph{pointers} allocated just before the kernel (in the same function) using
+\textit{malloc}.  With PGCC, the \texttt{\#pragma acc data} directive can be
+omitted; it can determine statically that 10,000 elements of $A$, $B$, and
+$C$ are accessed and thus need to be copied.  With Clang, if the
+\texttt{\#pragma omp target data} directive is omitted, the kernel crashes at
+runtime (\emph{Device Illegal Address}).
+
+\subsection{Translation Algorithm}
+
+\newcommand{\translate}[1]{\ensuremath{\mathcal{T}[\![ #1 ]\!]}}
+\begin{figure*}
+\begin{align*}
+%
+\translate{\texttt{\#pragma acc data\ }\textit{data-clauses}} &=
+   \texttt{\#pragma omp target data } \translate{\textit{data-clauses}} \\
+%
+\translate{\texttt{copyin(}\textit{vars}\texttt{)}} &=
+   \texttt{map(to:}\textit{vars}\texttt{)} \\
+%
+\translate{\texttt{copyout(}\textit{vars}\texttt{)}} &=
+   \texttt{map(from:}\textit{vars}\texttt{)} \\
+%
+\translate{\texttt{copy(}\textit{vars}\texttt{)}} &=
+   \texttt{map(tofrom:}\textit{vars}\texttt{)} \\
+%
+\translate{\texttt{create(}\textit{vars}\texttt{)}} &=
+   \texttt{map(alloc:}\textit{vars}\texttt{)} \\
+%
+\translate{\texttt{\#pragma acc parallel\ }\textit{parallel-clauses}} &=
+   \left.\parbox[t]{4in}{%
+   \texttt{\#pragma omp target teams\ }\translate{\textit{parallel-clauses}}}\right. \\
+%
+\translate{\texttt{num\_gangs(}\textit{n}\texttt{)}} &=
+   \texttt{num\_teams(}\textit{n}\texttt{)} \\
+%
+\translate{\texttt{vector\_length(}\textit{n}\texttt{)}} &=
+   \texttt{thread\_limit(}\textit{n}\texttt{)} \\
+%
+\translate{\texttt{private(}\textit{vars}\texttt{)}} &=
+   \texttt{private(}\textit{vars}\texttt{)} \\
+%
+\translate{\texttt{reduction(}\textit{vars}\texttt{)}} &=
+   \texttt{reduction(}\textit{reductions}\texttt{)} \\
+%
+\translate{\texttt{\#pragma acc loop\ }\textit{type}\texttt{\ }\textit{loop-clauses}} &=
+   \begin{cases}
+   \texttt{\#pragma omp distribute\ }\translate{\textit{loop-clauses}} \\
+        \hspace*{2em} \textup{ if $\textit{type} \neq \texttt{seq}$, and the loop is the outermost} \\
+        \hspace*{3em} \textup{ non-sequential loop in a loop nest} \\
+        \hspace*{3em} \textup{ in a \texttt{parallel} region} \vspace*{.25em} \\
+   \texttt{\#pragma omp parallel for\ }\translate{\textit{loop-clauses}} \\
+        \hspace*{2em} \textup{ if $\textit{type} \neq \texttt{seq}$, and the loop is the innermost} \\
+        \hspace*{3em} \textup{ non-sequential loop in a loop nest} \\
+        \hspace*{3em} \textup{ in a \texttt{parallel} region} \vspace*{.25em} \\
+   \texttt{\#pragma omp distribute parallel for\ }\translate{\textit{loop-clauses}} \\
+        \hspace*{2em} \textup{ if both of the above apply} \vspace*{.33em} \\
+   \epsilon
+        \hspace*{1.65em} \textup{ otherwise} \\
+   \end{cases} \\
+%
+\translate{\texttt{independent}} &=
+   \epsilon \\
+%
+\end{align*}
+\caption{Translation rules from OpenACC to OpenMP~4.  The function
+$\mathcal{T}$ maps phrases derived from the OpenACC subset grammar to
+strings corresponding to one or more OpenMP directives.  Sequences
+(e.g., \textit{parallel-clauses}) are translated elementwise, that is,
+$\translate{t_1\ t_2\ \dots\ t_n} =
+\translate{t_1}\ \translate{t_2}\ \dots\ \translate{t_n}$.}
+\label{fig:rules}
+\end{figure*}
+
+In the previous section, we showed that OpenACC and OpenMP~4 device directives
+handle work distribution differently, and that differences in compilers can
+have significant effects on the performance (and correctness) of code.  This
+means that porting from OpenACC to OpenMP~4 will almost certainly involve
+some manual effort.  Nevertheless, much of the translation is straightforward
+and tedious, which makes it worth automating.
+
+Our tool implements the translation rules in Figure~\ref{fig:rules}.  The
+majority of the translation is straightforward: Each phrase in OpenACC is
+mapped to a corresponding phrase in OpenMP\@.  These translation rules were
+designed with the following objectives:
+\begin{itemize}
+\item \emph{The translation should be straightfoward and consistent.}
+Each OpenACC directive should translate to the same OpenMP directive(s)
+every time, and the translation should mirror the structure of the OpenACC
+code, to the extent possible.
+\item \emph{Automated translation and manual editing should be complementary.}
+The tool should not try to automate too much.  It should not perform loop
+nest optimizations; it should not infer implicit data transfers; it should
+not make any invasive code changes.  The tool should automate the tedious,
+most obvious parts of the translation, in a completely predictable way.
+More sophisticated changes---like performance tuning---should be under the
+control of the programmer.
+\end{itemize}
+
+\subsubsection{Translating Data Directives}
+
+It is important to note that, according to the translation rules in
+Figure~\ref{fig:rules}, \emph{each data directive is translated as-is}.  No
+attempt is made to infer implicit data transfers or array lengths.
+
+\subsubsection{Translating Loop Nests}
+
+OpenACC's \texttt{gang} clause is roughly equivalent to OpenMP's
+\texttt{distribute}---effectively, it distributes work among thread blocks.
+Likewise, OpenACC's \texttt{vector} clause is roughly equivalent to OpenMP's
+\texttt{parallel for} inside a \texttt{teams} region: it distributes work among
+threads within a block.
+
+Unfortunately, translation isn't quite so simple.  OpenACC does not
+\emph{require} the \texttt{gang} and \texttt{vector} clauses to be used, in
+which case the work distribution is determined by the compiler.  OpenACC is
+also surprisingly lax about where the nesting of gang and vector loops.  For
+example, a \texttt{gang} loop may be nested inside a \texttt{vector} loop;
+there is no direct equivalent of this in OpenMP.  Moreover, PGCC may ignore the
+\texttt{gang} and \texttt{vector} directives altogether, e.g., when a gang loop
+appears inside another gang loop, or a vector loop appears outside another
+vector loop.
+
+For these reasons, \emph{the loop-handling rule of Figure~\ref{fig:rules}
+ignores OpenACC's \texttt{gang} and \texttt{vector} clauses}.  Instead, it uses
+a heuristic typical of parallelizing compilers (one that PGCC appears to use,
+in fact): the iterations of the outermost parallelizable loop are partitioned
+among thread blocks, and the iterations of the innermost parallelizable loop
+are partitioned among the CUDA threads within that block.
+
 \subsection{Translation Tool}
@@
 \section{Evaluation}
+
+PGCC~16.1 and Clang~3.8
+
 \todo{1 page including tables - evaluation: what code did we test it on; how many lines changed; what performance numbers did we get - Nawrin}

### 2016-04-28 sig-alternate-sample.tex
Added CUDA execution model

@@
 \section{Overview}
+\subsection{CUDA Execution Model}
+CUDA is parallel programming platform created by NVIDIA that allows programmer to write ``kernel"-- function executed on device by running many GPU threads concurrently.
+Only one kernel executes at a time.
+All of the threads in a kernel are collectively called a ``grid".
+A grid contains multiple thread blocks, organized into one, two, or three-dimensional array.
+Again, each thread block is organized into one, two, or three-dimensional array of threads.
+Each thread block is assigned to one Streaming Multiprocessor (SM).
+Each SM consists of several Streaming Processors (SPs).
+Each thread of a thread block is assigned to one of the SPs within the SM.
+In CUDA, thread blocks are typically divided into thread groups called Warps.
+Warps are executed in single instruction, multiple data (SIMD) fashion.
+The SM can context-switch from one wrap to another with zero-overhead thread scheduling.
+Again threads within a block can coordinate using barrier synchronization whereas threads in different blocks cannot (no execution order for block)~\cite{kirk2012programming}.
 \subsection{OpenACC Directives}

### 2016-04-29 sig-alternate-sample.tex
Sections 1 and 2

--- 2016-04-24 sig-alternate-sample.tex
@@
 \usepackage{xspace}
+\hyphenpenalty=750
 \newcommand\todo[1]{\textcolor{red}{** #1}}
@@
 % DOI
-\doi{10.475/123_4}
+\doi{}
 
 % ISBN
-\isbn{123-4567-24-567/08/06}
+\isbn{}
 
 %Conference
-\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}
+\conferenceinfo{XSEDE '16}{July 17--21, 2016, Miami, FL, USA}
 
@@
 % --- Author Metadata here ---
-\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
 %\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
@@
 
-\title{Toward Automatic Translation from OpenACC to OpenMP 4}
-%\subtitle{[Extended Abstract]
-%\titlenote{A full version of this paper is available as
-%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-%\LaTeX$2_\epsilon$\ and BibTeX} at
-%\texttt{www.acm.org/eaddress.htm}}}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
+\title{From OpenACC to OpenMP 4: \\ Toward Automatic Translation}
 
-\numberofauthors{8} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
+\def\sharedaffiliation{%
+\end{tabular}
+\begin{tabular}{c}}
+
+\numberofauthors{4}
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
 \alignauthor
-Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{1932 Wallamaloo Lane}\\
-       \affaddr{Wallamaloo, New Zealand}\\
-       \email{trovato@corporation.com}
-% 2nd. author
+Nawrin Sultana \\
+\email{nzs0034@auburn.edu}
 \alignauthor
-G.K.M. Tobin\titlenote{The secretary disavows
-any knowledge of this author's actions.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{P.O. Box 1212}\\
-       \affaddr{Dublin, Ohio 43017-6221}\\
-       \email{webmaster@marysville-ohio.com}
-% 3rd. author
-\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
-one who did all the really hard work.}\\
-       \affaddr{The Th{\o}rv{\"a}ld Group}\\
-       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
-       \affaddr{Hekla, Iceland}\\
-       \email{larst@affiliation.org}
+Alexander Calvert \\
+\email{afc0005@auburn.edu}
+\alignauthor
+Jeffrey L.\@ Overbey \\
+\email{joverbey@auburn.edu}
+\sharedaffiliation
+       \affaddr{Department of Computer Science and Software Engineering}\\
+       \affaddr{Auburn University}\\
+       \affaddr{Auburn University, AL, USA}
 \and  % use '\and' if you need 'another row' of author names
-% 4th. author
-\alignauthor Lawrence P. Leipuner\\
-       \affaddr{Brookhaven Laboratories}\\
-       \affaddr{Brookhaven National Lab}\\
-       \affaddr{P.O. Box 5000}\\
-       \email{lleipuner@researchlabs.org}
-% 5th. author
-\alignauthor Sean Fogarty\\
-       \affaddr{NASA Ames Research Center}\\
-       \affaddr{Moffett Field}\\
-       \affaddr{California 94035}\\
-       \email{fogartys@amesres.org}
-% 6th. author
-\alignauthor Charles Palmer\\
-       \affaddr{Palmer Research Laboratories}\\
-       \affaddr{8600 Datapoint Drive}\\
-       \affaddr{San Antonio, Texas 78229}\\
-       \email{cpalmer@prl.com}
+\alignauthor
+\alignauthor
+Galen Arnold \\
+\email{gwarnold@illinois.edu}
+\alignauthor
+\sharedaffiliation
+       \affaddr{National Center for Supercomputing Applications}\\
+       \affaddr{University of Illinois at Urbana-Champaign}\\
+       \affaddr{Urbana, IL, USA}
 }
-% There's nothing stopping you putting the seventh, eighth, etc.
-% author on the opening page (as the 'third row') but we ask,
-% for aesthetic reasons that you place these 'additional authors'
-% in the \additional authors block, viz.
-\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
-email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
-(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
-\date{30 July 1999}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
+
+\date{}
 
@@
 \section{Introduction}
-\todo{1 page - introduction/summary of the work - Dr. Overbey}
 
-\section{Motivation}
+While CUDA and OpenCL are the canonical APIs for programming accelerator
+devices, scientific programmers are increasingly turning to directive-based
+APIs, which are more concise and declarative.  Since it first appeared in 2011,
+OpenACC has been the standard directive-based API for scientific programmers
+targeting NVIDIA GPUs.  OpenMP~4.0 added directives for accelerators, inspired
+by OpenACC\@.  Both OpenACC and OpenMP are supported in both C/C++ and Fortran.
+
+OpenACC is supported by PGI's compilers.  Intel's compilers support OpenMP~4.
+Cray recently announced that their support for OpenACC will be frozen at
+version~2.0, and future development will focus on OpenMP~4 device directives.
+Recent versions of GCC support OpenACC; recent versions of Clang support
+OpenMP~4.
+
+So, the two APIs are rivals, at least for the time being.
+
+Since nearly every compiler supports OpenMP for multicore parallelism, it is
+reasonable to expect that the device directives added in OpenMP~4 will
+eventually supplant OpenACC\@.  However, that certainly has not happened
+yet---and given the strong backing for OpenACC from PGI/NVIDIA, there may
+continue to be some rivalry between the two specifications.
+
+In the XSEDE ecosystem, the PGI compilers are installed on Comet (at SDSC),
+allowing OpenACC to be used for programming its NVIDIA GPU nodes.  At TACC,
+portions of Stampede are being upgraded to the second-generation Intel Xeon Phi
+(codenamed ``Knight's Landing''), which can be programmed using OpenMP~4.
+
+Since OpenACC inspired the device directives added in OpenMP~4.0, the two APIs
+are quite similar.  In many cases, there is a 1--1 mapping between directives.
+However, there are some important differences as well.
+
+In this paper, we investigate the feasibility of a tool to semi-automatically
+convert OpenACC code to use OpenMP~4 device directives.  We have prototyped a
+tool that automates a large part of the conversion process; we have also
+identified parts of the conversion process that must be (or should be) left to
+the programmer.
+
+We focused on converting OpenACC to OpenMP~4 \emph{that targets the same
+accelerator device}---an NVIDIA GPU.  We used PGI~C/C++~16.1 as our OpenACC
+compiler, and we used Clang~3.8 as our OpenMP compiler.
+
+We chose this setup because it allowed us to investigate changes in semantics
+and performance that result from \emph{changing the compiler and API}, rather
+than those that result from a change in target hardware.  Admittedly, for XSEDE
+users, a more realistic scenario would be converting OpenACC on Comet (which
+targets NVIDIA GPUs) to OpenMP on Stampede (which targets Xeon Phi's).
+However, changing the target hardware in addition to the compiler and API
+introduces additional complexity.  Keeping the target hardware fixed allowed us
+to focus on the differences between OpenACC and OpenMP, rather than differences
+in the accelerator devices themselves.
+
+We used the EPCC Level~1 benchmarks as our test cases, along with a few
+miscellaneous examples, and we focused on the subset of OpenACC used by these
+programs.  Our results show that \emph{some} aspects of OpenACC-to-OpenMP
+conversion can be automated, but some manual intervention is essential.
+Also, both APIs rely heavily on the compiler, potentially making a change in
+compilers one of the biggest obstacles in converting code from OpenACC to
+OpenMP~4.
+
+The remainder of this paper is organized as follows.
+Section~\ref{sec:bg} gives a brief history and overview of OpenACC and OpenMP\@.
+Section~\ref{sec:translation} discusses translation from OpenACC to OpenMP,
+focusing on complications that prohibit completely automatic translation.
+Section~\ref{sec:algorithm} describes our tool and the algorithm it uses to
+convert OpenACC to OpenMP device directives.
+Section~\ref{sec:eval} presents empirical data describing the conversion.
+Section~\ref{sec:future} describes limitations of the current tool and
+directions for future work.
+
+\section{Background}
+\label{sec:bg}
+
 \subsection{History of OpenACC and OpenMP}
-\todo{Alexander}
-\subsection{Future of OpenACC and OpenMP}
-\todo{Alexander}
-\subsection{Relevance to XSEDE}
-\todo{Dr. Overbey}
 
-\section{Overview}
+OpenACC has a short history. It was originally developed by PGI, Cray, and
+NVIDIA, with version 1.0 being released in 2011. Version 2.0 was released in
+2013, adding support for procedure calls and nested parallelization (enabled by
+new CUDA capabilities), among other features. Version 2.5 was a more minor
+revision, made in 2015.
+
+%In the 1980's, SMP vendors typically defined directives for specifying how their SMP's divided up work on their individual processors. In the late 90's, 
+OpenMP dates back to the late 1990s, when the OpenMP Architecture Review Board
+was formed to develop a specification for directive-based parallelism that
+could be used commonly across multiple vendors' multiprocessors. The first
+OpenMP specification was released for Fortran in 1997 and for C/C++ in 1998.
+The OpenMP specification has undergone several revisions since that time.
+Notably, %task parallelism was added in version~3.0, and 
+version~4.0 added support for accelerators in 2013.  The so-called
+\emph{device directives} or \emph{target directives} added to support this
+were based on OpenACC.  However, as we will see, the two APIs are not
+identical.
+
+\subsection{CUDA Execution Model}
+
+OpenACC and OpenMP device directives are both designed to offload computations
+to accelerator devices---massively parallel devices like GPUs and Xeon Phis,
+designed for data parallel computations.  Each API has its own execution model,
+which is intentionally abstract to avoid coupling it to the specifics of any
+one device.
+
+For the purposes of this paper, we will only focus on one type of accelerator
+device: NVIDIA GPUs.  To understand the behavior of OpenACC and OpenMP
+constructs on these devices, it is helpful to understand their execution
+model---the CUDA execution model.
+
+CUDA-capable devices (i.e., NVIDIA GPUs) execute \emph{kernels}, functions that
+execute on the device by running many GPU threads concurrently. Only one kernel
+executes at a time. All of the threads in a kernel are collectively called a
+\emph{grid}. A grid is comprised of one or more \emph{thread blocks}, and each
+thread block is comprised of one or more \emph{threads}.
+
+In hardware, each thread block is assigned to one Streaming Multiprocessor
+(SM). Each SM consists of several Streaming Processors (SPs); each thread of a
+thread block is assigned to one of the SPs within the SM.  Thread blocks are
+divided into 32-thread groups called \emph{warps}.  Instructions are fetched
+and executed on a per-warp basis, not on a per-thread basis, so the 32 threads
+within each warp execute instructions in single instruction, multiple data
+(SIMD) fashion.  The SM context switches from one warp to another at every
+instruction issue (``zero-overhead thread scheduling''), so while threads
+within a warp execute in lockstep, threads in \emph{different} warps may not.
+Threads within a block can coordinate using barrier synchronization, whereas
+threads in different blocks cannot~\cite{kirk2012programming}.
+
+\subsection{OpenMP~4 Directives}
+
+\emph{Parallel.}
+OpenMP's fundamental construct is the parallel construct. When a parallel
+construct is encountered by a thread, that thread creates a \emph{team} of new
+threads, becoming that team's master. The threads are assigned numbers, with
+the master given number 0. The threads each execute a copy of the parallel
+region's code. %None of the child threads may continue execution beyond the
+%parallel region, and if any thread terminates within the region, the work done
+%up to that point is undefined. Nested parallelism is legal in OpenMP, so
+%parallel constructs may contain nested parallel constructs.
+
+\emph{For.}
+Inside a parallel region, a \emph{for} construct may be specified. A for region
+is associated with a for loop, and upon encountering the region, a different
+thread from the current team is assigned to each iteration of the for loop. A
+parallel for region is also defined as a shorthand for a parallel region
+containing nothing but a single for region. OpenMP 4.0 adds a similar
+construct, simd, which divides work of the associated loop up among SIMD lanes
+of the current thread.
+
+\emph{Target Data.}
+A target data region may be specified as well. A target data region creates a
+data environment for execution. It may specify a device clause, which specifies
+a device to execute the environment on, and a map clause, which specifies a
+mapping of data from the host to the device.
+
+\emph{Target Teams.}
+A target teams construct is equivalent to a target construct containing a teams
+construct. The target teams construct creates a data environment and a
+\emph{league} of thread teams to execute it.
+
+\emph{Distribute.}
+A distribute construct creates a league of teams and specifies that contained
+loops will be executed by those teams.
+
+\emph{Distribute Parallel For.}
+A distribute parallel for construct combines distribute, parallel, and for
+construct to specify that a loop can be executed in parallel by multiple
+threads in multiple teams.
+
 \subsection{OpenACC Directives}
-\todo{Alexander}
-\subsection{OpenMP 4 Directives}
-\todo{Alexander}
+
+OpenACC has fewer constructs. This is in part because OpenACC is specifically
+for use with massively parallel accelerator hardware, where OpenMP supports
+accelerators as well as multicore CPUs. %As a result, task parallelism is not
+%present in OpenACC.
+However, many of OpenACC's construct are analogous to OpenMP ones.
+
+\emph{Parallel.}
+OpenACC also has a parallel construct. A parallel construct, when encountered,
+creates \emph{gangs of workers} to execute the region on the accelerator.
+
+\emph{Loop.}
+A loop construct is also available, specifying the type of parallelism for the
+associated loop. There is also the parallel loop abbreviation, a shorthand for
+a loop construct immediately inside a parallel construct.
+
+\emph{Kernels.}
+The kernels construct specifies a region that is to be divided into a sequence
+of kernels for accelerator execution. The loop and shorthand kernels loop are
+available for kernels directives as they are for parallel directives.
+
+\emph{Data.}
+A data directive is also defined. It specifies for a region what data should be
+copied into the accelerator upon entry to the region and copied out on exit and
+defines what data is to be allocated on the device for the duration of the
+region. %Alternatively, the specification provides enter data and exit data
+%directives that specify data allocation in the same way, but without defining a
+%region -  enter data device allocations simply exist until exit data directives
+%deallocate them.
 
 \section{Translation}
-\todo{1 page including screenshot(s) - description of our tool: what directives it handles, how it works, where to get it - Nawrin}
+\label{sec:translation}
+
 The major compilers supporting OpenACC are Portland Group (PGI), Cray, and CAPS.
@@
 
-Our tool implements the translation rules in Figure~\ref{fig:rules}.  The
-majority of the translation is straightforward: Each phrase in OpenACC is
-mapped to a corresponding phrase in OpenMP\@.  These translation rules were
-designed with the following objectives:
+Our translation algorithm was designed with the following objectives:
 \begin{itemize}
@@
 
+The translation algorithm is as follows.  The input is a C source file
+containing OpenACC directives.  The output is the same file with these
+directives replaced by OpenMP directives.
+\begin{enumerate}
+\item If any of the following occur, warn the user; these structures will
+not be automatically converted:
+\begin{itemize}
+\item The header file \texttt{openacc.h} is included.
+\item The macro \texttt{\_OPENACC} is used.
+\item Functions in the OpenACC runtime library are called or referenced.
+\end{itemize}
+\item Label loops using the algorithm described in \S\ref{sec:loops}.
+\item Apply the translation rules in Figure~\ref{fig:rules} to convert
+OpenACC directives to OpenMP directives.
+\end{enumerate}
+
+Our tool implements the translation rules in Figure~\ref{fig:rules}.  The
+majority of the translation is straightforward: Each phrase in OpenACC is
+mapped to a corresponding phrase in OpenMP\@.
+
 \subsubsection{Translating Data Directives}
@@
 \subsubsection{Translating Loop Nests}
+\label{sec:loops}
 
@@
 \section{Evaluation}
+\label{sec:eval}
 
@@
 \section{Limitations and Future Work}
+\label{sec:future}
 \todo{1 page - limitations, including directives not handled, and future work - Dr. Overbey}

### 2016-04-29 sig-alternate-sample.tex
Section 3 and start of 4

@@
 
-\newcommand{\squishlist}{
-    \begin{list}{$\bullet$}
+\newcommand{\squish}{
      { \setlength{\itemsep}{0pt}      \setlength{\parsep}{3pt}
@@
 
-\section{Translation}
+\section{Translation Considerations}
 \label{sec:translation}
 
-The major compilers supporting OpenACC are Portland Group (PGI), Cray, and CAPS.
+%The major compilers supporting OpenACC are Portland Group (PGI), Cray, and CAPS.
 On the other hand, Intel has decided not to support OpenACC until accelerators are fully integrated into OpenMP.
@@
 
+Our goal was to create a tool to convert OpenACC code to equivalent OpenMP
+code, automating as much of the process as possible.  For the purposes of
+prototyping, we focused on translating C code (as opposed to C++ or Fortran).
+
 \subsection{A Subset of OpenACC}
@@
 OpenACC described by the grammar in Figure~\ref{fig:grammar}.  This includes
-all of the directives and clauses used by the EPCC Benchmarks, as well as the
-other example programs in our test suite.  While it omits some important
+all of the directives and clauses used by the example programs in our test
+suite (notably, the EPCC Level~1 benchmarks).  While it omits some important
 features---asynchronous execution, unstructured data lifetimes, etc.---it does
@@
 \textit{acc-directive} &$\rightarrow$& \texttt{\#pragma acc data\ }\textit{data-clauses} \\
-                       &     $|$     & \texttt{\#pragma acc parallel}\texttt{\ }\textit{parallel-clauses} \\
+                       &     $|$     & \texttt{\#pragma acc parallel}\texttt{\ }\textit{par-clauses} \\
                        &     $|$     & \texttt{\#pragma acc loop\ }\textit{type}\texttt{\ }\textit{loop-clauses} \\
-\textit{type}          &$\rightarrow$& \texttt{gang} \\
+\textit{type}          &$\rightarrow$& \texttt{gang vector} \\
+                       &     $|$     & \texttt{gang} \\
                        &     $|$     & \texttt{vector} \\
                        &     $|$     & \texttt{seq} \\
-                       &     $|$     & $\epsilon$ \\
+%                       &     $|$     & $\epsilon$ \\
 \textit{data-clause}   &$\rightarrow$& \texttt{copyin(}\textit{vars}\texttt{)} \\
@@
                        &     $|$     & \texttt{create(}\textit{vars}\texttt{)} \\
-\textit{parallel-clause}&$\rightarrow$&\texttt{num\_gangs(}\textit{n}\texttt{)} \\
+\textit{par-clause}&$\rightarrow$&\texttt{num\_gangs(}\textit{n}\texttt{)} \\
                        &     $|$     & \texttt{vector\_length(}\textit{n}\texttt{)} \\
@@
 \caption{OpenACC subset grammar accepted by the translation tool.  Syntax is
-omitted for \textit{vars}, \textit{reductions}, and integers \textit{n}.  The
-nonterminals \textit{data-clauses} and \textit{loop-clauses} correspond to
+omitted for \textit{vars}, \textit{reductions}, and natural numbers \textit{n}.
+The nonterminals \textit{data-clauses} and \textit{loop-clauses} correspond to
 sequences of zero or more phrases derived from \textit{data-clause} and
@@
 commercial compiler for OpenACC\@.  We compiled OpenMP~4 using a custom build
-of Clang~3.8\footnote{See \url{https://parallel-computing.pro/index.php/9-cuda/43-openmp-4-0-on-nvidia-cuda-gpus}} (PGCC does not yet support
-OpenMP~4\footnote{Cray's compilers support both, but there are currently no
-Cray resources in the XSEDE ecosystem.}).  As we will show, directives that
+of Clang~3.8~\cite{clang} (since PGCC does not yet support
+OpenMP~4).\footnote{Cray's compilers support both, but there are currently no
+Cray resources in the XSEDE ecosystem.}  As we will show, directives that
 appear similar may be translated quite differently by the two compilers.
 
-\subsubsection{Work Distribution Differs}
+\subsubsection{Work Distribution for Single Loops}
 
-\begin{figure*}
+\begin{figure}
 \begin{center}
@@
 \begin{verbatim}
-#pragma acc data copyin(A[0:N], B[0:N]), copyout(C[0:N])
-#pragma omp target data map(to:A[0:N], B[0:N]), map(from:C[0:N])
+#pragma acc data copyin(A[0:N], B[0:N]), \
+                 copyout(C[0:N])
+#pragma omp target data map(to:A[0:N], B[0:N]), \
+                        map(from:C[0:N])
 {
@@
 \end{center}
-\caption{Vector addition}
+\caption{Vector addition kernel.}
 \label{fig:vecadd}
-\end{figure*}
+\end{figure}
 
@@
 and thread blocks on the CUDA device.  When $N = 10000$:
-\begin{itemize}
+\begin{itemize}\squish
 \item PGCC generates a grid of 79 thread blocks with 128 threads per block;
@@
 \begin{figure*}
-\begin{minipage}{.5\textwidth}
+\begin{minipage}{.33\textwidth}
+\emph{OpenACC (Original):}
 {\footnotesize
 \begin{verbatim}
-// OpenACC
 #pragma acc parallel loop
-for (int i = 0; i < 10; i++) {      // Gang
-    for (int j = 0; j < 10; j++) {  // Vector
+for (int i = 0; i < 10; i++) {
+  for (int j = 0; j < 10; j++) {
     ...
@@
 }
+\centerline{(a)}
 \end{minipage}
-\begin{minipage}{.5\textwidth}
+\begin{minipage}{.33\textwidth}
+\emph{OpenACC (Equivalent):}
 {\footnotesize
 \begin{verbatim}
-// OpenMP 4 Device Directives
+#pragma acc parallel loop gang
+for (int i = 0; i < 10; i++) {
+  #pragma acc parallel loop vector
+  for (int j = 0; j < 10; j++) {
+    ...
+  }
+}
+\end{verbatim}
+}
+\centerline{(b)}
+\end{minipage}
+\begin{minipage}{.33\textwidth}
+\emph{OpenMP~4:}
+{\footnotesize
+\begin{verbatim}
 #pragma omp target teams distribute
@@
 }
+\centerline{(c)}
 \end{minipage}
@@
 
+\subsubsection{Work Distribution for Loop Nests}
+\label{sec:nests}
+
 More important is the handling of loop nests, such as the simple
-doubly nested loops in Figure~\ref{fig:nest}.
+doubly nested loops in Figure~\ref{fig:nest}(a).
 
@@
 innermost loop as a vector loop.  Given the OpenACC code in
-Figure~\ref{fig:nest}, the compiler does exactly this: the iterations of the
+Figure~\ref{fig:nest}(a), the compiler does exactly this: the iterations of the
 $i$-loop are divided among the thread blocks, and the iterations of the
-$j$-loop are divided among the threads within each block.
+$j$-loop are divided among the threads within each block.  The OpenACC code in
+Figure~\ref{fig:nest}(b) is functionally equivalent but makes this distribution
+explicit.
 
-In OpenMP, work distribution is not so automatic.  Consider these
-``obvious'' but incorrect translations of the OpenACC code into OpenMP: The
-\texttt{\#pragma acc parallel loop} directive could be replaced by
-\begin{itemize}
+In OpenMP, work distribution is not so automatic.  Consider these ``obvious''
+but incorrect translations: The \texttt{\#pragma acc parallel loop} directive
+could be replaced by:
+\begin{itemize}\squish
 \item \texttt{\#pragma omp target teams}.  In this case, the entire loop
@@
 parallelization directive is applied to only one loop.  OpenMP requires the
-programmer to explicitly specify how to run each loop in the nest.  The
-translation most similar to the OpenACC code is shown on the right in
-Figure~\ref{fig:nest}: The iterations of the $i$-loop are divided among thread
-blocks using the \texttt{distribute} directive, and the iterations of the
-$j$-loop are divided among threads in the block using the \texttt{parallel for}
-directive.
+programmer to explicitly specify how to run each loop in the nest.
+
+The translation most similar to the OpenACC code is shown in
+Figure~\ref{fig:nest}(c).  The iterations of the $i$-loop are divided among
+thread blocks using the \texttt{distribute} directive, and the iterations of
+the $j$-loop are divided among threads in the block using the \texttt{parallel
+for} directive.
+%
+OpenACC's \texttt{gang} clause is roughly equivalent to OpenMP's
+\texttt{distribute}---effectively, it distributes work among thread blocks.
+Likewise, OpenACC's \texttt{vector} clause is roughly equivalent to OpenMP's
+\texttt{parallel for} inside a \texttt{teams} region: it distributes work among
+threads within a block.
 
@@
 \texttt{\#pragma omp target data} directive is omitted, the kernel crashes at
-runtime (\emph{Device Illegal Address}).
+runtime (\emph{device illegal address}).
 
-\subsection{Translation Algorithm}
+\section{Translation Algorithm and \\ Tool Implementation}
+\label{sec:algorithm}
+
+In the previous section, we showed that OpenACC and OpenMP~4 device directives
+handle work distribution differently, and that differences in compilers can
+have significant effects on the performance (and correctness) of code.  This
+means that porting from OpenACC to OpenMP~4 will almost certainly involve some
+manual effort.  Nevertheless, much of the translation is straightforward and
+tedious, which makes it worth automating.
+
+Therefore, we designed our translation tool---and its underlying
+algorithm---with the following objectives:
+\begin{itemize}
+\item \emph{The translation should be straightfoward and consistent.}
+Each OpenACC directive should translate to the same OpenMP directive(s)
+every time, and the translation should mirror the structure of the OpenACC
+code, to the extent possible.
+\item \emph{Automated translation and manual editing should be complementary.}
+The tool should not try to automate too much.  It should not perform loop
+nest optimizations; it should not infer implicit data transfers; it should
+not make any invasive code changes.  The tool should automate the tedious,
+most obvious parts of the translation, in a completely predictable way.
+More sophisticated changes---like performance tuning---should be under the
+control of the programmer.
+\end{itemize}
+
+The translation algorithm is as follows.  The input is a C source file
+containing OpenACC directives.  The output is the same file with OpenMP
+directives substituted.
+\begin{enumerate}
+\item If any of the following occur, warn the user; these structures will
+not be automatically converted:
+\begin{itemize}
+\item The header file \texttt{openacc.h} is included.
+\item The macro \texttt{\_OPENACC} is used.
+\item Functions in the OpenACC runtime library are called or referenced.
+\end{itemize}
+\item Prepare parallel loop nests using the algorithm described in
+\S\ref{sec:loops} below.
+\item Apply the translation rules described below in \S\ref{sec:directives} to
+convert OpenACC directives to OpenMP directives.
+\end{enumerate}
+
+\subsection{Loop Nest Preparation}
+\label{sec:loops}
+
+As noted in \S\ref{sec:nests} above, OpenACC-to-OpenMP translation is
+complicated by the fact that work distribution is explicit in OpenMP, while it
+may be left to the compiler in OpenACC\@.  We noted that the code in
+Figure~\ref{fig:nest}(a) is equivalent to that in Figure~\ref{fig:nest}(b),
+expect the latter makes the work distribution explicit.
+
+Before our translation tool attempts to translate any OpenACC directives,
+it ``prepares'' each parallel loop nest as follows.
+\begin{itemize}\squish
+\item If the loop nest contains explicit gang and vector clauses, there must be
+exactly one gang loop and exactly one vector loop in the nest.  The same loop
+may be both the gang and vector loop.
+\item If the loop nest does not contain an explicit gang loop, the outermost
+non-sequential loop is chosen as the gang loop.
+\item If the loop nest does not contain an explicit vector loop, the innermost
+non-sequential loop is chosen as the vector loop.
+\item All loops in the loop nest other than the gang and vector loop are run
+sequentially.
+\item If a loop will be run sequentially, and it contains a \texttt{private}
+clause, all \texttt{private} variables must not be live at the top of the loop.
+\item If the loop nest does not have the form above (e.g., if it contains two
+gang loops, or if a vector loop appears outside a gang loop), the translation
+to OpenMP cannot proceed.
+\end{itemize}
+
+The last rule is necessary because OpenACC is surprisingly lax about where the
+nesting of gang and vector loops.  For example, a \texttt{gang} loop may be
+nested inside a \texttt{vector} loop; there is no direct equivalent of this in
+OpenMP.  Moreover, PGCC may ignore the \texttt{gang} and \texttt{vector}
+directives altogether, e.g., when a gang loop appears inside another gang loop,
+or a vector loop appears outside another vector loop.
+
+The next-to-last rule ensures that, if a loop is run sequentially, the
+\texttt{private} clause can be removed without changing the behavior of the
+loop.  Although the rule is stated in terms of a live variables
+analysis~\cite{kennedy}, the concept is straightforward.  If a variable $a$
+is declared as private, but each iteration assigns a value to $a$ before
+reading its value, then the \texttt{private} clause can be eliminated if
+the loop is run sequentially.  It cannot be removed if each iteration of the
+loop depends on the value of $a$ prior to the loop (e.g., if the first
+instruction in the loop was \texttt{a++}).
+
+The second and third rules implement a heuristic typical of parallelizing
+compilers (one that PGCC appears to use, in fact): the iterations of the
+outermost parallelizable loop are partitioned among thread blocks, and the
+iterations of the innermost parallelizable loop are partitioned among the CUDA
+threads within that block.
+
+To determine if a loop is parallelizable (for the second and third rules), our
+tool uses a combination of syntactic checks and dependence
+analysis~\cite{wolfe,kennedy}:
+\begin{enumerate}\squish
+\item If the loop contains a \texttt{seq} clause, it is not parallelizable.
+\item If the loop contains a \texttt{gang} or \texttt{vector} clause, it is
+parallelizable.
+\item Otherwise, we perform a dependence analysis on the loop.  If the loop
+does not carry a dependence, it is parallelizable~\cite{kennedy}; if it
+does, we determine it to be non-parallelizable.
+\end{enumerate}
+
+\subsection{Directive Translation}
+\label{sec:directives}
 
@@
 %
-\translate{\texttt{\#pragma acc parallel\ }\textit{parallel-clauses}} &=
+\translate{\texttt{\#pragma acc parallel\ }\textit{par-clauses}} &=
    \left.\parbox[t]{4in}{%
-   \texttt{\#pragma omp target teams\ }\translate{\textit{parallel-clauses}}}\right. \\
+   \texttt{\#pragma omp target teams\ }\translate{\textit{par-clauses}}}\right. \\
 %
@@
    \begin{cases}
+   \texttt{\#pragma omp distribute parallel for\ }\translate{\textit{loop-clauses}} \\
+        \hspace*{2em} \textup{ if $\textit{type} = \texttt{gang vector}$} \vspace*{.25em} \\
    \texttt{\#pragma omp distribute\ }\translate{\textit{loop-clauses}} \\
-        \hspace*{2em} \textup{ if $\textit{type} \neq \texttt{seq}$, and the loop is the outermost} \\
-        \hspace*{3em} \textup{ non-sequential loop in a loop nest} \\
-        \hspace*{3em} \textup{ in a \texttt{parallel} region} \vspace*{.25em} \\
+        \hspace*{2em} \textup{ if $\textit{type} = \texttt{gang}$} \vspace*{.25em} \\
    \texttt{\#pragma omp parallel for\ }\translate{\textit{loop-clauses}} \\
-        \hspace*{2em} \textup{ if $\textit{type} \neq \texttt{seq}$, and the loop is the innermost} \\
-        \hspace*{3em} \textup{ non-sequential loop in a loop nest} \\
-        \hspace*{3em} \textup{ in a \texttt{parallel} region} \vspace*{.25em} \\
-   \texttt{\#pragma omp distribute parallel for\ }\translate{\textit{loop-clauses}} \\
-        \hspace*{2em} \textup{ if both of the above apply} \vspace*{.33em} \\
+        \hspace*{2em} \textup{ if $\textit{type} = \texttt{vector}$} \vspace*{.25em} \\
    \epsilon
-        \hspace*{1.65em} \textup{ otherwise} \\
+        \hspace*{1.65em} \textup{ if $\textit{type} = \texttt{seq}$} \\
    \end{cases} \\
@@
 strings corresponding to one or more OpenMP directives.  Sequences
-(e.g., \textit{parallel-clauses}) are translated elementwise, that is,
+(e.g., \textit{par-clauses}) are translated elementwise, that is,
 $\translate{t_1\ t_2\ \dots\ t_n} =
 \translate{t_1}\ \translate{t_2}\ \dots\ \translate{t_n}$.}
+The symbol $\epsilon$ denotes the empty string.
 \label{fig:rules}
@@
 
-In the previous section, we showed that OpenACC and OpenMP~4 device directives
-handle work distribution differently, and that differences in compilers can
-have significant effects on the performance (and correctness) of code.  This
-means that porting from OpenACC to OpenMP~4 will almost certainly involve
-some manual effort.  Nevertheless, much of the translation is straightforward
-and tedious, which makes it worth automating.
-
-Our translation algorithm was designed with the following objectives:
-\begin{itemize}
-\item \emph{The translation should be straightfoward and consistent.}
-Each OpenACC directive should translate to the same OpenMP directive(s)
-every time, and the translation should mirror the structure of the OpenACC
-code, to the extent possible.
-\item \emph{Automated translation and manual editing should be complementary.}
-The tool should not try to automate too much.  It should not perform loop
-nest optimizations; it should not infer implicit data transfers; it should
-not make any invasive code changes.  The tool should automate the tedious,
-most obvious parts of the translation, in a completely predictable way.
-More sophisticated changes---like performance tuning---should be under the
-control of the programmer.
-\end{itemize}
-
-The translation algorithm is as follows.  The input is a C source file
-containing OpenACC directives.  The output is the same file with these
-directives replaced by OpenMP directives.
-\begin{enumerate}
-\item If any of the following occur, warn the user; these structures will
-not be automatically converted:
-\begin{itemize}
-\item The header file \texttt{openacc.h} is included.
-\item The macro \texttt{\_OPENACC} is used.
-\item Functions in the OpenACC runtime library are called or referenced.
-\end{itemize}
-\item Label loops using the algorithm described in \S\ref{sec:loops}.
-\item Apply the translation rules in Figure~\ref{fig:rules} to convert
-OpenACC directives to OpenMP directives.
-\end{enumerate}
+The loop nest preparation algorithm in the previous section identifies a single
+gang loop and a single vector loop in each parallel loop nest, and it guarantees
+that the vector loop is nested under (or is the same as) the gang loop.  After
+gang, vector, and sequential loops have been identified, it is possible to
+proceed with translating OpenACC directives to OpenMP directives.
 
@@
 
-\subsubsection{Translating Data Directives}
-
 It is important to note that, according to the translation rules in
@@
 
-\subsubsection{Translating Loop Nests}
-\label{sec:loops}
-
-OpenACC's \texttt{gang} clause is roughly equivalent to OpenMP's
-\texttt{distribute}---effectively, it distributes work among thread blocks.
-Likewise, OpenACC's \texttt{vector} clause is roughly equivalent to OpenMP's
-\texttt{parallel for} inside a \texttt{teams} region: it distributes work among
-threads within a block.
-
-Unfortunately, translation isn't quite so simple.  OpenACC does not
-\emph{require} the \texttt{gang} and \texttt{vector} clauses to be used, in
-which case the work distribution is determined by the compiler.  OpenACC is
-also surprisingly lax about where the nesting of gang and vector loops.  For
-example, a \texttt{gang} loop may be nested inside a \texttt{vector} loop;
-there is no direct equivalent of this in OpenMP.  Moreover, PGCC may ignore the
-\texttt{gang} and \texttt{vector} directives altogether, e.g., when a gang loop
-appears inside another gang loop, or a vector loop appears outside another
-vector loop.
-
-For these reasons, \emph{the loop-handling rule of Figure~\ref{fig:rules}
-ignores OpenACC's \texttt{gang} and \texttt{vector} clauses}.  Instead, it uses
-a heuristic typical of parallelizing compilers (one that PGCC appears to use,
-in fact): the iterations of the outermost parallelizable loop are partitioned
-among thread blocks, and the iterations of the innermost parallelizable loop
-are partitioned among the CUDA threads within that block.
+Also, these rules assume that each loop in a parallel loop nest has been
+\emph{explicitly} labeled with a \texttt{gang}, \texttt{vector}, or
+\texttt{seq} clause (or both \texttt{gang} and \texttt{vector}).  Of course,
+this is not necessary.  In fact, our tool runs the loop nest preparation
+algorithm implicitly; it identifies a gang and vector loop in each parallel
+loop nest, but it does not literally insert OpenACC \texttt{gang},
+\texttt{vector}, and \texttt{seq} clauses into the source code (although
+it retains them if they are already present in the code).
 

### 2016-04-29 sig-alternate-sample.tex
Evaluation, etc.

@@
 \item Functions in the OpenACC runtime library are called or referenced.
+\item The \texttt{kernels} directive is used.
 \end{itemize}
@@
 \end{enumerate}
+
+The \texttt{\#pragma acc kernels} directive is effectively a request for
+automatic parallelization; the compiler must decide what loops to parallelize
+and how.  There is no direct equivalent in OpenMP\@.  Thus, our algorithm does
+not support the kernels directive.  Instead, the programmer should modify the
+code to explicitly parallelize loops using the \texttt{parallel} directive.
 
@@
 \subsection{Translation Tool}
-We prototyped the refactoring in a PLDT OpenACC Eclipse plug-in (C analyses and refactorings for OpenACC).
-Our tool support various data flow analyses, dependence analysis, refactorings, and parser for OpenACC directives.
-It provides an user interface (UI) for refactorings.
-Other than UI, it also supports command line interface (CLI), so one can run the refactorings outside Eclipse.
 
-Our tool does not handle all the OpenACC directives.
+We prototyped our OpenACC-to-OpenMP translation algorithm by extending the
+Eclipse C/C++ Development Tools.  In addition to our translation algorithm, we
+added support for various data flow analyses and dependence analysis, along
+with a parser for OpenACC directives.  Our translation can be performed from
+within the Eclipse user interface, but we also added a command line interface
+(CLI), so one can run the translation from the command line, without installing
+Eclipse.
+
+%Our tool does not handle all the OpenACC directives.
 We transformed the directives present in EPCC level 1 OpenACC benchmark except ``kernel" directive.
@@
 Then analyses and source translations are executed.
-
-\subsection{Porting OpenACC To OpenMP Directives (POOD)}
-The \pood refactoring converts OpenACC directives to OpenMP 4.0 directives.
-
-\textbf{Motivation:} While porting OpenACC to OpenMP it is necessary for directives to be compatible. Without automated tooling, a programmer has to manually transform each directives.
-
-\textbf{Precondition:} A programmer selects a C source file and invokes the \pooda transformation. The following precondition is checked:
-
-\begin{itemize}
-\item Source file contains OpenACC directives.
-\item OpenACC directives in source code supported by our transformation tool.
-\end{itemize}
-
-\textbf{Mechanism:} The refactoring introduces OpenMP directives that correspond to OpenACC directives.
-
-POOD first determines all the preprocessor include statements and check whether there is any ``openacc.h".
+%  
+%\subsection{Porting OpenACC To OpenMP Directives (POOD)}
+%The \pood refactoring converts OpenACC directives to OpenMP 4.0 directives.
+%
+%\textbf{Motivation:} While porting OpenACC to OpenMP it is necessary for directives to be compatible. Without automated tooling, a programmer has to manually transform each directives.
+%
+%\textbf{Precondition:} A programmer selects a C source file and invokes the \pooda transformation. The following precondition is checked:
+%
+%\begin{itemize}
+%\item Source file contains OpenACC directives.
+%\item OpenACC directives in source code supported by our transformation tool.
+%\end{itemize}  
+%
+%\textbf{Mechanism:} The refactoring introduces OpenMP directives that correspond to OpenACC directives.
+%
+%POOD first determines all the preprocessor include statements and check whether there is any ``openacc.h".
 If openacc header presents POOD removes it.\todo{need to remove openacc functions}.
@@
 When all the transformations have completed, POOD writes the changes to file.
+%
+%
+%\begin{table}[t]
+%	\begin{minipage}{0.47\textwidth}
+%%\small
+%\caption{Equivqlent OpenACC and OpenMP clause}
+%\centering
+%\begin{tabular}{ | l@{\hspace{0.1cm}}|  @{\hspace{0.4cm}}r@{\hspace{0.1cm}}|} 
+% \hline
+% \hspace*{0.4cm} OpenACC Clause & OpenMP Clause\hspace*{0.6cm}\\
+% \hspace*{0.5cm} (\#pragma acc)%\footnote{The ID is used to refer to each application in the rest of the paper.} 
+%  & (\#pragma omp)\hspace*{0.6cm} \\
+% \hline
+% data & target data \\ 
+% copy(A[0:n*n]) & map(tofrom:A[0:n*n]) \\
+% copyin(A[0:n*n]) & map(to:A[0:n*n]) \\
+% copyout(A[0:n*n]) & map(from:A[0:n*n]) \\
+% create(A[0:n*n]) & map(to:A[0:n*n]) \\ 
+% parallel & target/teams \\ 
+% parallel loop private(t1,t2) & parallel for private(t1,t2) \\
+% loop & for \\
+% loop private(tmp) & for private(tmp) \\
+% loop reduction(+:t1,t2) & for reduction(+:t1,t2) \\
+% loop independent & for independent \\
+% parallel loop & target/teams/distribute \\
+% \hline
+%\end{tabular}
+%
+%\label{table:1}
+%\end{minipage}
+%\end{table}
+%%\normalsize
+%
+%
+%\textbf{Example:}
 
-\begin{table}[t]
-	\begin{minipage}{0.47\textwidth}
-%\small
-\caption{Equivqlent OpenACC and OpenMP clause}
-\centering
-\begin{tabular}{ | l@{\hspace{0.1cm}}|  @{\hspace{0.4cm}}r@{\hspace{0.1cm}}|} 
- \hline
- \hspace*{0.4cm} OpenACC Clause & OpenMP Clause\hspace*{0.6cm}\\
- \hspace*{0.5cm} (\#pragma acc)%\footnote{The ID is used to refer to each application in the rest of the paper.} 
-  & (\#pragma omp)\hspace*{0.6cm} \\
- \hline
- data & target data \\ 
- copy(A[0:n*n]) & map(tofrom:A[0:n*n]) \\
- copyin(A[0:n*n]) & map(to:A[0:n*n]) \\
- copyout(A[0:n*n]) & map(from:A[0:n*n]) \\
- create(A[0:n*n]) & map(to:A[0:n*n]) \\ 
- parallel & target/teams \\ 
- parallel loop private(t1,t2) & parallel for private(t1,t2) \\
- loop & for \\
- loop private(tmp) & for private(tmp) \\
- loop reduction(+:t1,t2) & for reduction(+:t1,t2) \\
- loop independent & for independent \\
- parallel loop & target/teams/distribute \\
- \hline
-\end{tabular}
-
-\label{table:1}
-\end{minipage}
-\end{table}
-%\normalsize
-
-\textbf{Example:}
-
-\todo{0.5 page - description of the conversion algorithm (what clauses translate to what?) - Nawrin/Dr.Overbey} 
 \section{Evaluation}
@@
 
-PGCC~16.1 and Clang~3.8
+%\begin{table}
+%\begin{tabular}{ | l@{\hspace{0.1cm}}|  @{\hspace{0.4cm}}r@{\hspace{0.1cm}}|} 
+%Test Code    & Number of Directives \\
+%EPCC Level~1 & \\
+%
+%\end{tabular}
+%\end{table}
 
-\todo{1 page including tables - evaluation: what code did we test it on; how many lines changed; what performance numbers did we get - Nawrin}
-\subsection{Test Corpus}
-\subsection{Performance}
+To evaluate the correctness of our tool, we used it to translate the EPCC
+Level~1 OpenACC benchmarks~\cite{epcc}, which are ports of the PolyBench and
+PolyBench/GPU kernels~\cite{polybench}.  We also translated several
+miscellaneous test programs, including naive and tiled matrix multiplication
+kernels.  We compiled the original OpenACC code using PGCC~16.1 and the
+resulting OpenMP code using Clang~3.8.  Performance measurements were taken on
+a Tesla K40 (Comet and Stampede have Tesla K20 GPUs).
+
+The EPCC Level~1 benchmarks contain a total of 70 OpenACC directives, three of
+which are \emph{kernels} directives that our tool would not translate directly.
+Of the remaining 67 directives, 13 were data directives, 23 were parallel loop
+directives, and 31 were loop directives nested under other parallel directives.
+There was only one OpenACC API call: a call to \texttt{acc\_init}, used to
+initialize the GPU and the OpenACC runtime.
+
+This suggests that a tool like ours could be quite valuable.  Only two changes
+needed to be made by the programmer: removing \texttt{acc\_init} and converting
+the three \emph{kernels} regions.  The remaining 67 directives could all be
+converted automatically, without manual intervention.
 
@@
 \label{sec:future}
-\todo{1 page - limitations, including directives not handled, and future work - Dr. Overbey}
 
-\section{Related Work}
-\section{Conclusion}
-\todo{Dr. Overbey}
+Our tool is a prototype, and our translation algorithm supports only a few
+commonly used OpenACC directives.  Some areas for improvement are obvious.  The
+algorithm needs to be expanded to include the entire OpenACC specification,
+including asynchronous kernel launches, unstructured data lifetimes, and
+OpenACC API calls.  Our tool works only on C, while OpenACC and OpenMP both
+support C++ and Fortran.  While the EPCC benchmarks were useful for evaluating
+our prototype, we need to test its effectiveness on larger code bases (i.e.,
+application codes), ideally by working with other XSEDE collaborators.
+
+While automating a tedious 1--1 translation is helpful, it is not the end of
+the conversion process.  As noted earlier, a common XSEDE use case will likely
+be converting OpenACC code to take advantage of the forthcoming
+second-generation Intel Xeon Phis on Stampede.  Our translation algorithm (and
+tool) could certainly be specialized for this process.
+
+After the initial OpenMP code has been created, some restructuring will almost
+certainly be necessary.  For example, we have performed some preliminary
+experiments with Intel's compilers on first-generation Phis; in one case, we
+needed to convert global variables to local variables to ensure that the OpenMP
+\emph{target data} directive would guarantee a copy of the arrays to the Phi.
+Cataloging (and perhaps automating) such transformations would be beneficial.
+
+Perhaps more importantly, performance portability is not guaranteed.  A direct
+translation of OpenACC code optimized for an NVIDIA GPU will not necessary
+perform at its peak on a Xeon Phi.
+
+Interestingly, we did not even obtain comparable performance targeting the
+\emph{same} hardware.  In our test cases, the GPU kernels generated by Clang
+were almost always several times slower than those generated by PGCC.  For
+example, consider the naive matrix multiplication kernel in
+Figure~\ref{fig:matmul}.  With $2048 \times 2048$ matrices, the OpenACC kernel
+ran for about 250~ms; the OpenMP kernel ran for about 925~ms.  Since the
+OpenACC and OpenMP codes were nearly identical, and they were running on the
+same hardware, we did not expect such massive differences in performance.
+Clearly, details of the compiler-generated code can have a substantial impact
+on performance.  Again, documenting differences in compilers would be
+helpful---what sort of performance change is typical, what can be done to
+improve performance when changing from compiler X to compiler Y, etc.
+
+\begin{figure}
+\begin{verbatim}
+#pragma omp target teams distribute
+#pragma acc parallel loop gang
+for (int j = 0; j < N; j++) {
+    #pragma omp parallel for
+    #pragma acc loop vector
+    for (int i = 0; i < N; i++) {
+        double t = 0.0;
+        #pragma acc loop seq
+        for (int k = 0; k < N; k++) {
+            t += m[k][i] * n[j][k];
+        }
+        p[j][i] = t;
+    }
+}
+\end{verbatim}
+\caption{Naive matrix multiplication.}
+\label{fig:matmul}
+\end{figure}
+
+%\section{Conclusion}
 
 %ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
+%\section{Acknowledgments}
 

### 2016-04-29 paper.tex
Renamed to paper.tex
Abstract

--- 2016-04-29 sig-alternate-sample.tex
@@
 \begin{abstract}
-
+For the past few years, OpenACC has been the primary directive-based API for
+programming accelerator devices like GPUs.  OpenMP~4.0 is now a competitor in
+this space, with support from different vendors.  In this paper, we describe an
+algorithm to convert (a subset of) OpenACC to OpenMP~4; we implemented this
+algorithm in a prototype tool and evaluated it by translating the EPCC Level~1
+OpenACC benchmarks.  We discuss some of the challenges in the conversion
+process and propose what parts of the process should be automated, what should
+be done manually by the programmer, and what future research and development is
+necessary in this area.
 \end{abstract}
@@
 \bibliographystyle{abbrv}
-\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
+\bibliography{paper}
 % You must have a proper ".bib" file

### 2016-04-29 paper.tex
Wrote conclusion

@@
 
-Our tool is a prototype, and our translation algorithm supports only a few
+Our tool is a prototype, and our translation algorithm supports only the most
 commonly used OpenACC directives.  Some areas for improvement are obvious.  The
@@
 
-%\section{Conclusion}
+\section{Conclusion}
+
+We have described an algorithm (\S\ref{sec:algorithm}) and a prototype tool
+that converts a subset of OpenACC (Figure~\ref{fig:grammar}) to OpenMP~4.  The
+translation algorithm was designed to be simple and predicatble, since it will
+necessarily be used in conjunction with manual restructuring and performance
+tuning.  We evaluated our tool by converting the EPCC Level~1 OpenACC
+benchmarks from OpenACC compiled by PGI C/C++ to OpenMP~4 compiled by Clang.
+Both targeted an NVIDIA GPU (Tesla~K40).  To keep the translation process
+transparent, the programmer was required to manually convert OpenACC
+\emph{kernels} loops, although the tool was able to automatically translate the
+remaining 67 of 70 OpenACC directives
 
@@
 %\section{Acknowledgments}
+
+\newpage
 

### 2016-04-29 paper.tex
CCS categories

@@
 <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+<concept_id>10010147.10010169.10010175</concept_id>
+<concept_desc>Computing methodologies~Parallel programming languages</concept_desc>
 <concept_significance>500</concept_significance>
@@
 <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
+<concept_id>10010147.10011777.10011014</concept_id>
+<concept_desc>Computing methodologies~Concurrent programming languages</concept_desc>
 <concept_significance>300</concept_significance>
@@
 <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
+<concept_id>10011007.10011006.10011073</concept_id>
+<concept_desc>Software and its engineering~Software maintenance tools</concept_desc>
+<concept_significance>500</concept_significance>
 </concept>
@@
 
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
+\ccsdesc[500]{Computing methodologies~Parallel programming languages}
+\ccsdesc[500]{Software and its engineering~Software maintenance tools}
 
@@
 
-\keywords{ACM proceedings; \LaTeX; text tagging}
+\keywords{accelerators; GPUs; OpenACC; OpenMP; translation}
 
@@
 OpenACC's \texttt{gang} clause is roughly equivalent to OpenMP's
-\texttt{distribute}---effectively, it distributes work among thread blocks.
+\texttt{distribute}: effectively, it distributes work among thread blocks.
 Likewise, OpenACC's \texttt{vector} clause is roughly equivalent to OpenMP's
@@
 %
+\translate{\texttt{independent}} &=
+   \epsilon \\
+%
 \translate{\texttt{\#pragma acc loop\ }\textit{type}\texttt{\ }\textit{loop-clauses}} &=
@@
    \end{cases} \\
-%
-\translate{\texttt{independent}} &=
-   \epsilon \\
 %
@@
 \emph{kernels} loops, although the tool was able to automatically translate the
-remaining 67 of 70 OpenACC directives
+remaining 67 of 70 directives.
 

### 2016-04-29 paper.tex
Add table with execution times

@@
 \documentclass{sig-alternate-05-2015}
-\usepackage{xcolor}
+\usepackage[T1]{fontenc}
+\usepackage[latin1]{inputenc}
 \usepackage{xspace}
+\usepackage[table]{xcolor}
 \hyphenpenalty=750
@@
 PolyBench/GPU kernels~\cite{polybench}.  We also translated several
-miscellaneous test programs, including naive and tiled matrix multiplication
+miscellaneous test programs, including na\"{i}ve and tiled matrix multiplication
 kernels.  We compiled the original OpenACC code using PGCC~16.1 and the
@@
 
-\section{Limitations and Future Work}
-\label{sec:future}
-
-Our tool is a prototype, and our translation algorithm supports only the most
-commonly used OpenACC directives.  Some areas for improvement are obvious.  The
-algorithm needs to be expanded to include the entire OpenACC specification,
-including asynchronous kernel launches, unstructured data lifetimes, and
-OpenACC API calls.  Our tool works only on C, while OpenACC and OpenMP both
-support C++ and Fortran.  While the EPCC benchmarks were useful for evaluating
-our prototype, we need to test its effectiveness on larger code bases (i.e.,
-application codes), ideally by working with other XSEDE collaborators.
-
-While automating a tedious 1--1 translation is helpful, it is not the end of
-the conversion process.  As noted earlier, a common XSEDE use case will likely
-be converting OpenACC code to take advantage of the forthcoming
-second-generation Intel Xeon Phis on Stampede.  Our translation algorithm (and
-tool) could certainly be specialized for this process.
-
-After the initial OpenMP code has been created, some restructuring will almost
-certainly be necessary.  For example, we have performed some preliminary
-experiments with Intel's compilers on first-generation Phis; in one case, we
-needed to convert global variables to local variables to ensure that the OpenMP
-\emph{target data} directive would guarantee a copy of the arrays to the Phi.
-Cataloging (and perhaps automating) such transformations would be beneficial.
-
-Perhaps more importantly, performance portability is not guaranteed.  A direct
-translation of OpenACC code optimized for an NVIDIA GPU will not necessary
-perform at its peak on a Xeon Phi.
-
-Interestingly, we did not even obtain comparable performance targeting the
-\emph{same} hardware.  In our test cases, the GPU kernels generated by Clang
-were almost always several times slower than those generated by PGCC.  For
-example, consider the naive matrix multiplication kernel in
-Figure~\ref{fig:matmul}.  With $2048 \times 2048$ matrices, the OpenACC kernel
-ran for about 250~ms; the OpenMP kernel ran for about 925~ms.  Since the
-OpenACC and OpenMP codes were nearly identical, and they were running on the
-same hardware, we did not expect such massive differences in performance.
-Clearly, details of the compiler-generated code can have a substantial impact
-on performance.  Again, documenting differences in compilers would be
-helpful---what sort of performance change is typical, what can be done to
-improve performance when changing from compiler X to compiler Y, etc.
+Interestingly, while both OpenACC and Clang were targeting the same GPU device,
+the GPU kernels generated by Clang were almost always slower than those
+generated by PGCC.  The naive matrix multiplication kernel in
+Figure~\ref{fig:matmul} can serve as a small, concrete example.  With $2048
+\times 2048$ matrices, the OpenACC kernel ran for about 250~ms; the OpenMP
+kernel ran for about 925~ms.  Since the OpenACC and OpenMP codes were nearly
+identical, and they were running on the same hardware, we did not expect such
+massive differences in performance.  Execution times for the EPCC Level~1
+benchmarks on a Tesla~K40 are shown in Table~\ref{table:epcc}.  The benchmarks
+were run with a data size of 1024 and 5 repetitions.  Clearly, details of the
+compiler-generated code can have a substantial impact on performance.
+%Again, documenting differences in compilers would be helpful---what sort of
+%performance change is typical, what can be done to improve performance when
+%changing from compiler X to compiler Y, etc.
 
@@
 \end{verbatim}
-\caption{Naive matrix multiplication.}
+\caption{Na\"{i}ve matrix multiplication.}
 \label{fig:matmul}
 \end{figure}
+
+\begin{table}
+{\sf
+\rowcolors{0}{gray!25}{white}
+\begin{tabular}{l@{\hspace*{1em}}r@{\hspace*{1em}}r@{\hspace*{1em}}r}
+\textbf{Kernel} & \textbf{PGI/OpenACC} & \textbf{Clang/OpenMP} & \textbf{Speedup} \\
+3MM     & 17241 &  50073 & 0.34 \\
+ATAX    & 24539 &  28308 & 0.87 \\
+BICG    & 25178 &  29561 & 0.85 \\
+MVT     & 24021 &  27928 & 0.86 \\
+SYRK    & 72562 & 213693 & 0.34  \\
+COV     & 28120 &  19905 & 1.41 \\
+COR     & 28640 & 187218 & 0.15 \\
+SYR2K   & 15231 &  66935 & 0.23 \\
+GESUMMV & 37998 &  52765 & 0.72 \\
+\multicolumn{4}{r}{\textit{Median: 0.72}}
+\end{tabular}
+}
+\caption{Execution times for EPCC Level~1 benchmarks on a Tesla~K40 ($\mu s$).}
+\label{table:epcc}
+\end{table}
+
+\section{Limitations and Future Work}
+\label{sec:future}
+
+Our tool is a prototype, and our translation algorithm supports only the most
+commonly used OpenACC directives.  Some areas for improvement are obvious.  The
+algorithm needs to be expanded to include the entire OpenACC specification,
+including asynchronous kernel launches, unstructured data lifetimes, and
+OpenACC API calls.  Our tool works only on C, while OpenACC and OpenMP both
+support C++ and Fortran.  While the EPCC benchmarks were useful for evaluating
+our prototype, we need to test its effectiveness on larger code bases (i.e.,
+application codes), ideally by working with other XSEDE collaborators.
+
+While automating a tedious 1--1 translation is helpful, it is not the end of
+the conversion process.  As noted earlier, a common XSEDE use case will likely
+be converting OpenACC code to take advantage of the forthcoming
+second-generation Intel Xeon Phis on Stampede.  Our translation algorithm (and
+tool) could certainly be specialized for this process.
+
+After the initial OpenMP code has been created, some restructuring will almost
+certainly be necessary.  For example, we have performed some preliminary
+experiments with Intel's compilers on first-generation Phis; in one case, we
+needed to convert global variables to local variables to ensure that the OpenMP
+\emph{target data} directive would guarantee a copy of the arrays to the Phi.
+Cataloging (and perhaps automating) such transformations would be beneficial.
+
+Perhaps more importantly, performance portability is not guaranteed.  A direct
+translation of OpenACC code optimized for an NVIDIA GPU will not necessary
+perform at its peak on a Xeon Phi.  As noted above, we did not even obtain
+comparable performance targeting the
+\emph{same} hardware.
 

### 2016-04-29 paper.tex
Start Related Work (Hernandez talk)

@@
 
+\section{Related Work}
+
+Most related to our work is a talk by Oscar Hernandez et al.\@ at
+OpenMPCon~2015~\cite{hernandez,hernandezvideo}.  This talk describes the
+salient differences between OpenACC~2.0 and OpenMP~4.0 and suggests a
+(manual) procedure for porting code from OpenACC to OpenMP\@.  Their procedure
+consists of five steps: removing constructs with no OpenMP counterparts (like
+\emph{kernels}), translating data regions, translating data updates, translating
+accelerator parallel regions, and adjusting function attribute specifiers.
+
 \section{Conclusion}
@@
 
-\newpage
+%\newpage
 

### 2016-04-29 paper.tex
Related work

@@
 
-Most related to our work is a talk by Oscar Hernandez et al.\@ at
+Most related to our work is a talk by Hernandez et al.\@ at
 OpenMPCon~2015~\cite{hernandez,hernandezvideo}.  This talk describes the
@@
 accelerator parallel regions, and adjusting function attribute specifiers.
+
+%Several directive-based GPU programming models have been proposed over the year to characterize parallel program. 
+Lee and Vetter~\cite{lee2012early} evaluate existing directive-based models and
+show that directive-based models can achieve reasonable performance compared to
+hand-written GPU codes. For accelarator programming, the two most promising
+options are OpenACC and OpenMP.  Wienke et al.~\cite{wienke2014pattern} compare
+both models with respect to their programmability to assist developers in
+deciding which approach to take.  Xu and colleagues~\cite{xu2013exploring}
+demonstrate the effectiveness of a hybrid model combined with OpenACC and
+OpenMP as a plausible solution to port scientific applications in heterogeneous
+architectures.
+
+%In our work we describe the context for porting OpenACC device directives to OpenMP directives. And proposed an automated tool (with a small manual intervention) to port OpenACC into OpenMP.
+
+\newpage
 
@@
 
-%\newpage
-
 %

### 2016-06-09 paper.tex
Added ACM copyright text
Fixed spelling errors

@@
 % Copyright
-\setcopyright{acmcopyright}
+\CopyrightYear{2016}
+\setcopyright{acmlicensed}
+\conferenceinfo{XSEDE16,}{July 17 - 21, 2016, , }
+\isbn{978-1-4503-4755-6/16/07}\acmPrice{\$15.00}
+\doi{http://dx.doi.org/10.1145/2949550.2949654}
+%\setcopyright{acmcopyright}
 %\setcopyright{acmlicensed}
@@
 % DOI
-\doi{}
+%\doi{}
 
 % ISBN
-\isbn{}
+%\isbn{}
 
 %Conference
-\conferenceinfo{XSEDE '16}{July 17--21, 2016, Miami, FL, USA}
-
-\acmPrice{\$15.00}
+%\conferenceinfo{XSEDE '16}{July 17--21, 2016, Miami, FL, USA}
+
+%\acmPrice{\$15.00}
 
@@
 show that directive-based models can achieve reasonable performance compared to
-hand-written GPU codes. For accelarator programming, the two most promising
+hand-written GPU codes. For accelerator programming, the two most promising
 options are OpenACC and OpenMP.  Wienke et al.~\cite{wienke2014pattern} compare
@@
 that converts a subset of OpenACC (Figure~\ref{fig:grammar}) to OpenMP~4.  The
-translation algorithm was designed to be simple and predicatble, since it will
+translation algorithm was designed to be simple and predictable, since it will
 necessarily be used in conjunction with manual restructuring and performance

### 2016-06-13 paper.tex
Add 4.2 Sequential Private/Reduction Removal

@@
 \usepackage[table]{xcolor}
+\usepackage{balance} % for \balance
 \hyphenpenalty=750
@@
 \S\ref{sec:loops} below.
+\item Remove \emph{private} and \emph{reduction} clauses from sequential
+loops as described in \S\ref{sec:privred} below.
 \item Apply the translation rules described below in \S\ref{sec:directives} to
@@
 \end{enumerate}
+
+\subsection{Sequential Private/Reduction Removal}
+\label{sec:privred}
+
+\begin{figure*}[t]
+\begin{center}
+{\footnotesize
+% (a) Original
+\begin{minipage}{.5\textwidth}
+\texttt{%
+int x, sum = 100; \\
+\textcolor{blue}{\bfseries \#pragma acc loop seq private(x) reduction(+:sum)} \\
+for (int i = 0; i < 5; i++) \{ \\
+\hspace*{2em} x = 100; \\
+\hspace*{2em} sum += x; \\
+\}
+}
+\end{minipage}
+\hspace*{.07\textwidth}
+% (b) Original
+\begin{minipage}{.36\textwidth}
+\texttt{%
+int x; \\
+\textcolor{blue}{\bfseries \#pragma acc loop seq reduction(+:x)} \\
+for (int i = 0; i < 5; i++) \\
+\hspace*{2em} x = 1;
+}
+\end{minipage} \\
+% Arrows
+\begin{minipage}{.5\textwidth}
+\centerline{\rule[-0.5em]{2em}{0em} $\Downarrow$}
+\end{minipage}
+\hspace*{.07\textwidth}
+\begin{minipage}{.36\textwidth}
+\centerline{\rule[-0.5em]{2em}{0em} $\Downarrow$}
+\end{minipage} \\
+% (a) Transformed
+\begin{minipage}{.5\textwidth}
+\texttt{%
+int x, sum = 100; \\
+for (int i = 0; i < 5; i++) \{ \\
+\hspace*{2em} \textcolor{blue}{\bfseries int x;} \\
+\hspace*{2em} x = 100; \\
+\hspace*{2em} sum += x; \\
+\}
+}
+\end{minipage}
+\hspace*{.07\textwidth}
+% (b) Transformed
+\begin{minipage}{.36\textwidth}
+\texttt{%
+int x; \\
+\textcolor{blue}{\bfseries int sum = 0;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (1)} \\
+for (int i = 0; i < 5; i++) \{ \\
+\hspace*{2em} \textcolor{blue}{\bfseries int x = 0;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (2)} \\
+\hspace*{2em} x = 1; \\
+\hspace*{2em} \textcolor{blue}{\bfseries sum += x;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (3)} \\
+\} \\
+\textcolor{blue}{\bfseries x = sum;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (4)}
+}
+\end{minipage} \\
+% Labels
+\begin{minipage}{.5\textwidth}
+\centerline{\rule[-0.5em]{2em}{0em} (a)}
+\end{minipage}
+\hspace*{.07\textwidth}
+\begin{minipage}{.36\textwidth}
+\centerline{\rule[-0.5em]{2em}{0em} (b)}
+\end{minipage} \\
+}
+\end{center}
+\caption{Elimination of \emph{private} and \emph{reduction} clauses on a sequential loop.}
+\label{fig:privred}
+\end{figure*}
+
+After a loop nest has been prepared, there will be a gang loop, a vector loop
+(possibly the same loop), and zero or more sequential loops.  OpenACC has a
+directive for sequential loops (\texttt{acc loop seq}), which permits
+\emph{private} and \emph{reduction} clauses.  However, OpenMP does not have a
+sequential loop directive.  This means that these directives---and more
+importantly, \emph{private} and \emph{reduction} clauses---must be removed from
+sequential loops in a way that will preserve the semantics of the
+\emph{private} and \emph{reduction} clauses.
+
+To remove a \emph{private} clause, we can simply re-declare the privatized
+variable in the loop body, as shown in Figure~\ref{fig:privred}(a) (noting that
+this features requires C99 support, which is available in all modern C
+compilers).  This declaration shadows the declaration in the outer scope,
+which gives the intended semantics of the \emph{private} clause.
+
+To illustrate removing a \emph{reduction} clause, consider as a concrete example
+the clause \texttt{reduction(+:x)}.  There are two cases.
+
+\textit{Case~1}.  This is perhaps the most commonly occurring use of the
+\emph{reduction} clause: If the only accesses to \texttt{x} in the loop
+body are in statements of the form \texttt{x++} or \texttt{x\ +=\
+}\emph{expression} (where \texttt{x} is not assigned in \emph{expression}),
+then the reduction clause can be removed with no change in semantics.  This is
+the case in Figure~\ref{fig:privred}(a), for example.
+
+\textit{Case~2}.  When the above condition does not hold, it is still possible
+to remove a \emph{reduction} clause.  However, preserving the semantics of the
+\emph{reduction} clause requires a more complex source code transformation,
+illustrated in Figure~\ref{fig:privred}(b).
+(1)~A ``fresh'' variable (\texttt{sum}) is declared to accumulate the result.
+(2)~Since \texttt{x} is the reduction variable, a local variable \texttt{x}
+is declared in the loop body with the same type as the shadowed variable
+\texttt{x}, and it is initialized to the initialization value for a sum
+reduction (\texttt{0}).
+(3)~At the end of the loop body, the local \texttt{x} is added to the result
+(\texttt{sum}).
+(4)~After the loop, the accumulated result is stored in the original
+\texttt{x}.
 
@@
 
-\newpage
-
 \section{Conclusion}
@@
 
+\balance
 %

### 2016-06-14 paper.tex
Fix reduction code; add reduction to loop-clauses

@@
                        &     $|$     & \texttt{private(}\textit{vars}\texttt{)} \\
+                       &     $|$     & \texttt{reduction(}\textit{reductions}\texttt{)} \\
 \end{tabular}
@@
 \texttt{%
-int x, sum = 100; \\
-\textcolor{blue}{\bfseries \#pragma acc loop seq private(x) reduction(+:sum)} \\
+int p, x = 100; \\
+\textcolor{blue}{\bfseries \#pragma acc loop seq private(p) reduction(+:x)} \\
 for (int i = 0; i < 5; i++) \{ \\
-\hspace*{2em} x = 100; \\
-\hspace*{2em} sum += x; \\
+\hspace*{2em} p = 100; \\
+\hspace*{2em} x += p; \\
 \}
@@
 \texttt{%
-int x; \\
+int x = 100; \\
 \textcolor{blue}{\bfseries \#pragma acc loop seq reduction(+:x)} \\
@@
 \texttt{%
-int x, sum = 100; \\
+int p, x = 100; \\
 for (int i = 0; i < 5; i++) \{ \\
-\hspace*{2em} \textcolor{blue}{\bfseries int x;} \\
-\hspace*{2em} x = 100; \\
-\hspace*{2em} sum += x; \\
+\hspace*{2em} \textcolor{blue}{\bfseries int p;} \\
+\hspace*{2em} p = 100; \\
+\hspace*{2em} x += p; \\
 \}
@@
 \texttt{%
-int x; \\
+int x = 100; \\
 \textcolor{blue}{\bfseries int sum = 0;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (1)} \\
@@
 \} \\
-\textcolor{blue}{\bfseries x = sum;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (4)}
+\textcolor{blue}{\bfseries x += sum;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (4)}
 }
@@
 To illustrate removing a \emph{reduction} clause, consider as a concrete example
-the clause \texttt{reduction(+:x)}.  There are two cases.
+\texttt{reduction(+:x)}.  There are two cases.
 
@@
 body are in statements of the form \texttt{x++} or \texttt{x\ +=\
-}\emph{expression} (where \texttt{x} is not assigned in \emph{expression}),
+}\emph{expression} (where \texttt{x} does not occur in \emph{expression}),
 then the reduction clause can be removed with no change in semantics.  This is
@@
 is declared in the loop body with the same type as the shadowed variable
-\texttt{x}, and it is initialized to the initialization value for a sum
+\texttt{x}, and it is initialized to the initializer value for a sum
 reduction (\texttt{0}).
@@
 (\texttt{sum}).
-(4)~After the loop, the accumulated result is stored in the original
+(4)~After the loop, the accumulated result is added to the original
 \texttt{x}.
